---
title: "Central America Portfolio -- BMZ/IKI Project Overview"
author: "Andreas Petutschnig, Leonie Gallé"
date: "2023-11-22"
output: workflowr::wflow_html
editor_options:
  chunk_output_type: console
---

```{r setup, message = FALSE, warning = FALSE, include = FALSE}
# load libraries and set session options. 
library(RColorBrewer)
library(dplyr)
library(knitr)
library(htmltools)
library(leaflet)
library(leaflet.extras)
library(leaflet.extras2)
library(leafpop)
library(magrittr)
library(mapme.biodiversity)
library(plotly)
library(reshape2)
library(scales)
library(sf)
options(scipen=10000)
```

```{r basemap_custom, message = FALSE, warning = FALSE, include = FALSE}
basemap_custom <-
  leaflet() %>%
  # add external map providers
  addTiles(group = "OpenStreetMap") %>%
  addProviderTiles(providers$CartoDB.Positron, group="CartoDB") %>%
  addProviderTiles(providers$Esri.WorldImagery, group="Satellite") %>%
  addProviderTiles(providers$Esri.WorldShadedRelief, group="Topography") %>%
  addProviderTiles(providers$NASAGIBS.ViirsEarthAtNight2012, group="Nightlights") %>%
  addFullscreenControl() %>%
  # add legend(s)
  addLayersControl(
     baseGroups = c("CartoDB", "OpenStreetMap", "Satellite", "Topography", "Nightlights"),
     options = layersControlOptions(collapsed = FALSE))
```


# Introduction

The following analysis intended to support a strategic peer-discussion. The goal of this exercise is to assist KfW and its partners to having a coherent overview of the Protected
Areas (short: PA) in Central America that are currently supported by KfW (Financial Cooperation). Additionally, the goal is to identify synergies in the different projects. The
analysis uses the publicly available data from the World Database of Protected Areas (WDPA/IUCN) [^1] as outlines.

# Visualized Protected Areas

We look at the Central America Portfolio (bilateral and regional, 11 projects). Data was downloaded from the WDPA and provided by the Project Partners. The following map shows the
Central America Portfolio and the PAs currently supported by KfW (Financial Cooperation). Each project has been assigned its own color. PAs which are supported in the scope of more
than one project are outlined in black.

# Map

This map is interactive, meaning that you can zoom into the map to see specific countries in more detail. Furthermore, supported PAs from the current portfolio of KfW are displayed
with their actual polygon boundaries. For analysis purposes it is useful to also activate the satellite layer which can often indicate whether areas are affected by agricultural
conversion.

```{r load_data, message = FALSE, warning = FALSE, include = FALSE}
dat_portfolio <- st_read("data/central_america_portfolio_with_bmz_numbers.gpkg")

dat_portfolio <- dat_portfolio %>%
  rename(
    `WDPA ID` = wdpa_id,
    `Project Name` = Project.Name,
    `Short Name` = Short.Name,
    `BMZ/IKI Number` = BMZ.Number,
    `PA Name` = PA.Name,
    `Project Start` = Project.Start,
    `Project End` = Project.End,
    `Size [ha]` = Size..ha.
  )

tbl_wdpas <- table(dat_portfolio$`WDPA ID`)
wdpa_duplicates <- names(tbl_wdpas) [tbl_wdpas > 1]

bmz_duplicates <- dat_portfolio %>%
  filter(`WDPA ID` %in% wdpa_duplicates) %>%
  select(`WDPA ID`, geom) %>%
  unique()

bmz_projects <- unique(dat_portfolio$`BMZ/IKI Number`)
n_bmz_projects <- length(bmz_projects)
col <- brewer.pal(n = n_bmz_projects, name = "Set3")

group_names <- as.character(bmz_projects)
for (i in seq_along(bmz_projects)) {
  prefix <- "BMZ Project"
  if (bmz_projects [i] == "209811209") {
    prefix <- "IKI Project"
  }
  group_names [i] <- paste(prefix, bmz_projects [i])
}

map_portfolio <- basemap_custom
map_portfolio <- map_portfolio %>%
    addPolygons(data = bmz_duplicates, fillOpacity = 0, opacity = 1, group = "BMZ Duplicates", color = "#000000", weight = 5)
for(i in seq_len(n_bmz_projects)) {
  polys <- dat_portfolio %>%
    filter(`BMZ/IKI Number` == bmz_projects[i])
  poly_color <- col [i]
  group_name <- group_names[i]
  
  map_portfolio <- map_portfolio %>%
    addPolygons(data = polys, fillOpacity = 0.9, fillColor = poly_color, group = group_name, color = "#666666",
                label = ~htmlEscape(`Project Name`), weight = 1, popup = popupTable(st_drop_geometry(polys)))
}

map_portfolio <- map_portfolio %>%
  addLayersControl(
     baseGroups = c("CartoDB", "OpenStreetMap", "Satellite", "Topography", "Nightlights"),
     overlayGroups = c(group_names),
     options = layersControlOptions(collapsed = FALSE)) %>%
  htmlwidgets::onRender("
        function() {
            $('.leaflet-control-layers-overlays').prepend('<label style=\"text-align:center\"><b>BMZ/IKI Project Overview</b></label>');
        }
  ")
```

```{r map_portfolio_plot, message = FALSE, warning = FALSE, echo = FALSE}
map_portfolio
```

To select/deselect individual BMZ Projects, you can use the layer overview on the right of the map.
The project numbers of overlapping BMZ projects are:

```{r duplicate-table, message = FALSE, warning = FALSE, echo = FALSE}
duplicate_table <- dat_portfolio %>%
  st_drop_geometry() %>%
  filter(`WDPA ID` %in% wdpa_duplicates) %>%
  select(`WDPA ID`, `BMZ/IKI Number`, `Short Name`) %>%
  unique() %>%
  arrange(., `WDPA ID`)

kable(
  duplicate_table
)
```


[comment]: <> (Literature)

[^1]: UNEP-WCMC and IUCN (2022), Protected Planet: The World Database on Protected Areas (WDPA) and World Database on Other Effective Area-based Conservation Measures (WD-OECM) [Online], February 2022, Cambridge, UK: UNEP-WCMC and IUCN. Available at: www.protectedplanet.net.
