---
title: "IUCN Action Types"
author: "Yota Eilers"
date: '2023-05-02'
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(warning = FALSE, message = FALSE, echo = FALSE)

# Load packages
library(readxl)
library(tidyverse)
library(stargazer)
library(reshape2)
library(knitr)
library(kableExtra)
```

## Data Description

## IUCN Action Types

IUCN suggests the following action type classification (note: descriptions are taken verbatim from IUCN):

**Land/Water Protection (1):** Actions to identify, establish or expand protected areas.

*1.1 Site/Area Protection*: Establishing or expanding public or private parks, reserves, and other protected areas roughly equivalent to IUCN Categories I-VI.

**Land/Water Management (2):** Actions directed at conserving or restoring sites, habitats and the wider environment.

**Species Management (3):** Actions directed at managing or restoring species, focused on the species of concern itself.

**Education & Awareness (4):** Actions directed at people to improve understanding and skills, and influence behavior.

**Law & Policy (5):** Actions to develop, change, influence and help implement formal legislation, regulation and voluntary standards.

**Livelihood, Economic & Other Incentives (6):** Actions to use economic and other incentives to influence behavior.

**External Capacity Building (7):** Actions to build the infrastructure to do better conservation.

```{r get list of bmz numbers}
# # list of panel datasets
# all_years <- list.files("/datadrive/datalake/mapme.protectedareas/output/tabular/regression_input/CEM/tn_projects") %>% 
#   grep("cem_tn_matched_panel_*", ., value = TRUE)
# bmz_list <- all_years %>% 
#     gsub("cem_tn_matched_panel_", "", .) %>% 
#     gsub(".csv", "", .)
# 
# bmz_list <- bmz_list[-which(bmz_list=="199566829"| bmz_list=="199966649")] # skip entries that cause issues because of perfect collinearity of treatment variable
# bmz_list <- bmz_list[-which(bmz_list=="201066661")] #skip because of coastal and marine protection without any forest cover
# # load each list entry and create unique panel id across all matching frames

iucnat <- read_excel("/datadrive/yota/newfiles/data/iucnat.xlsx",
                   sheet = "classification")
bmz_list <- unique(iucnat$bmz)
```


```{r}
# kfw_data <- read_excel("/datadrive/yota/test/Geodatamodel_KfW_extended_biodiversiy_KW_07.xlsx",
#                           sheet = "fill-me (ALT)", range ="A2:BG198")

kfw_data <- read_excel("/datadrive/yota/newfiles/data/iucnat_master.xlsx",
                       sheet = "fill-me (ALT)", skip = 1,
                       col_types = c(rep("guess", 49), rep("numeric", 10)))

# Select relevant columns
actiontypes <- kfw_data[,c(3,26:59)]
names(actiontypes)[1] <- "id"

# Cleaning the data set: rm cols where all NA
not_all_na <- function(x) any(!is.na(x))
actiontypes <- actiontypes %>% 
  select(where(not_all_na))

actiontypes <- actiontypes %>% 
  unite(col = "bmz", starts_with("PN"), sep = "", na.rm = TRUE) %>% 
  rename_with(~paste0("class", .), c(-id, -bmz)) %>% 
  filter(bmz %in% bmz_list)

at_reduced_long <- actiontypes %>% 
  pivot_longer(c(-id, -bmz)) %>% 
  mutate(name = str_extract(name, "class[0-9]")) %>% 
  group_by(bmz, name) %>%
  summarise(max = max(value))

at_reduced <- at_reduced_long %>% 
  pivot_wider(names_from = "name", values_from = "max",
              values_fill = 0) %>% 
  ungroup()

```

## By BMZ number

```{r, warning=FALSE, message=FALSE, echo=FALSE}
# By BMZ
at_bybmz <- actiontypes %>% 
  pivot_longer(c(-id, -bmz)) %>% 
  mutate(name = str_extract(name, "class[0-9]")) %>% 
  group_by(bmz, name) %>% 
  summarise(max = max(value)) %>% 
  group_by(max, name) %>% 
  summarise(count = n()) %>% 
  pivot_wider(names_from = "name", values_fill = 0, 
              values_from = "count") %>% 
  mutate(max = recode(max, 
                      "0" = "No focus",
                      "1" = "Weak focus",
                      "2" = "Strong focus",
                      .default = "Not specified"))

names(at_bybmz) <- c("Focus", "(1) Protection", "(2) Management", "(3) Species",
                 "(4) Education", "(5) Law/policy", "(6) Livelihood", "(7) Capacity")

# stargazer(at_bybmz,
#           type = "text",
#           summary = FALSE,
#           rownames = FALSE,
#           title = "IUCN action types: focus by project (BMZ) number")

at_bybmz %>%  
  kable()
```

## By Protected Area

```{r echo=FALSE, message=FALSE, warning=FALSE}
# By PA
at_bypa <- actiontypes %>% 
  pivot_longer(c(-id, -bmz)) %>% 
  mutate(name = str_extract(name, "class[0-9]")) %>% 
  group_by(id, name) %>% 
  summarise(max = max(value)) %>% 
  group_by(max, name) %>% 
  summarise(count = n()) %>% 
  pivot_wider(names_from = "name",  values_fill = 0,
              values_from = "count") %>% 
  mutate(max = recode(max, 
                      "0" = "No focus",
                      "1" = "Weak focus",
                      "2" = "Strong focus",
                      .default = "Not specified"))

names(at_bypa) <- c("Focus", "(1) Protection", "(2) Management", "(3) Species",
                     "(4) Education", "(5) Law/policy", "(6) Livelihood", "(7) Capacity")

# stargazer(at_bypa,
#           type = "text",
#           summary = FALSE,
#           rownames = FALSE,
#           title = "IUCN action types: focus by PA identifier") 
at_bypa %>%  
  kable()
```

## By BMZ and PA

```{r echo=FALSE, message=FALSE, warning=FALSE}
at_bypa2 <- at_bypa %>%  
  rename("F" = "Focus", "(1)" = "(1) Protection",
         "(2)" = "(2) Management", "(3)" = "(3) Species",
         "(4)" = "(4) Education", "(5)" = "(5) Law/policy", 
         "(6)" = "(6) Livelihood", "(7)" = "(7) Capacity")
new <- cbind(at_bybmz, at_bypa2) 
at_multicol <- new %>%  
  select("Focus", "(1) Protection", "(1)", "(2) Management", "(2)", "(3) Species", "(3)", 
         "(4) Education", "(4)", "(5) Law/policy", "(5)", 
         "(6) Livelihood", "(6)", "(7) Capacity", "(7)")

at_multicol %>%  
  kable(col.names = c("Focus by", "BMZ", "PA", "BMZ", "PA", "BMZ", "PA", "BMZ", "PA", 
                      "BMZ", "PA", "BMZ", "PA", "BMZ", "PA")) %>%
  add_header_above(c(" " = 1, "(1) Protection" = 2,  "(2) Management" = 2, "(3) Species" = 2,
                     "(4) Education" = 2, "(5) Law/Policy" = 2, "(6) Livelihood" = 2,
                     "(7) Capacity" = 2))

at_multicol %>%  
  kable(col.names = c("Focus by", "BMZ", "PA", "BMZ", "PA", "BMZ", "PA", "BMZ", "PA", 
                      "BMZ", "PA", "BMZ", "PA", "BMZ", "PA"),
        format = "latex") %>%
  add_header_above(c(" " = 1, "(1) Protection" = 2,  "(2) Management" = 2, "(3) Species" = 2,
                     "(4) Education" = 2, "(5) Law/Policy" = 2, "(6) Livelihood" = 2,
                     "(7) Capacity" = 2)) %>%
  save_kable("actiontypes_focus.tex")

```



## Overlap of projects by IUCN action type categories

Note that focus values were re-coded to a binary focus indicator. Only *strong focus* is coded as "focus" in this case, whereas *weak focus* or *no focus* are coded as "no focus", i.e. zero.

### By main categories

```{r}
# Recode focus such that strong focus is "1" and rest is "0"
at_recoded <- at_reduced_long %>% 
  mutate(max = recode(max,
                      `2` = 1,
                      `1` = 0,
                      `0` = 0)) %>% 
  pivot_wider(names_from = "name", values_from = "max",
              values_fill = 0) %>% 
  ungroup() %>%  
  na.exclude()

# Function for overlap of two vectors
get_overlap <- function(x, y) {
  # sum gives number of entries that match between the two vectors
  sum((x - y) == 0) / nrow(x)
}

# Function for 'overlap' matrix
cross_overlap <- function(start, end, data) {
  # Go through rows (inner map) and columns (outer mapö)
  tbl <- map_dfc(start:end, function(j) map_dbl(start:end, function(i) get_overlap(data[, i], data[, j])))
  names(tbl) <- names(data[, start:end])
  tbl
}

at_overlap <- as.matrix(cross_overlap(2, 8, at_recoded))

rownames(at_overlap) <- c("(1) Protection", "(2) Management", "(3) Species",
                     "(4) Education", "(5) Law/policy", "(6) Livelihood", "(7) Capacity")
colnames(at_overlap) <- c("(1) Protection", "(2) Management", "(3) Species",
                     "(4) Education", "(5) Law/policy", "(6) Livelihood", "(7) Capacity")

at_overlap_plot <- at_overlap

# Remove digits and lower triangle (redundant)
at_overlap <- at_overlap %>%
  format(digits = 2)
at_overlap[lower.tri(at_overlap)] <- ""
at_overlap %>%
  kable()

# Save latex table
kable(at_overlap, format = "latex") %>%
  save_kable("actiontypes_overlap.tex")

# Build plot
at_overlap_plot[lower.tri(at_overlap_plot)] <- NA 

at_overlap_plot %>% 
  melt(na.rm = TRUE) %>% 
  ggplot(aes(Var2, Var1, fill = value)) +
  geom_tile(color = "white")+
 scale_fill_gradient2(low = "blue", high = "red", mid = "white", 
   midpoint = 0, limit = c(0,1), space = "Lab", 
   name="Focus\nOverlap") +
  theme_minimal()+ 
 theme(axis.text.x = element_text(angle = 45, vjust = 1, size = 10, hjust = 1),
       axis.text.y = element_text(size = 10)) +
  xlab("") + 
  ylab("") +
 coord_fixed()

```


### By subcategories

```{r}
# Recode focus such that strong focus is "1" and rest is "0"
at_recoded_subg <- actiontypes %>% 
  mutate(across(starts_with("class"), 
                ~ recode(.,
                      `2` = 1,
                      `1` = 0,
                      `0` = 0))) %>% 
  filter(id != 5002) # this PA does not have any data

at_overlap_subg <- as.matrix(cross_overlap(2, 25, at_recoded_subg))

rownames(at_overlap_subg) <- names(at_recoded_subg[, 2:25])
colnames(at_overlap_subg) <- names(at_recoded_subg[, 2:25])

at_overlap_subg_plot <- at_overlap_subg

# Summary table for each class

# -- by PA
at_subg_bmz <- actiontypes %>% 
  group_by(bmz) %>% 
  summarise(across(starts_with("class"), ~max(.x)))

saveRDS(at_subg_bmz, "at_subg_bmz.RDS")

at_subg_tab_bmz <- apply(at_subg_bmz[-c(1)], 2, table)
maxlen <- max(sapply(at_subg_tab_bmz,length))
fillchar <- 0
t(sapply(at_subg_tab_bmz, function(x) c(x, rep(fillchar, maxlen - length(x)))))

# -- by BMZ
at_subg_tab <- apply(actiontypes[-c(1, 26)], 2, table)
maxlen <- max(sapply(at_subg_tab,length))
fillchar <- 0
t(sapply(at_subg_tab, function(x) c(x, rep(fillchar, maxlen - length(x)))))

# Remove digits and lower triangle (redundant)
at_overlap_subg <- at_overlap_subg %>%
  format(digits = 2)
at_overlap_subg[lower.tri(at_overlap_subg)] <- ""
at_overlap_subg %>%
  kable()

# Save latex table
kable(at_overlap_subg, format = "latex") %>%
  save_kable("actiontypes_overlap_subg.tex")

# Build plot
at_overlap_subg_plot[lower.tri(at_overlap_subg_plot)] <- NA 

at_overlap_subg_plot %>% 
  melt(na.rm = TRUE) %>% 
  ggplot(aes(Var2, Var1, fill = value)) +
  geom_tile(color = "white")+
 scale_fill_gradient2(low = "blue", high = "red", mid = "white", 
   midpoint = 0, limit = c(0,1), space = "Lab", 
   name="Focus\nOverlap") +
  theme_minimal()+ 
 theme(axis.text.x = element_text(angle = 45, vjust = 1, size = 10, hjust = 1),
       axis.text.y = element_text(size = 10)) +
  xlab("") + 
  ylab("") +
 coord_fixed()
```


```{r eval = FALSE}
# Recode focus such that strong focus is "1" and rest is "0"
at_cor <- at_reduced_long %>% 
  mutate(max = recode(max,
                      `2` = 1,
                      `1` = 0,
                      `0` = 0)) %>% 
  pivot_wider(names_from = "name", values_from = "max",
              values_fill = 0) %>% 
  ungroup() %>%  
  select(-bmz) %>% 
  na.exclude() %>% 
  cor()

rownames(at_cor) <- c("(1) Protection", "(2) Management", "(3) Species",
                     "(4) Education", "(5) Law/policy", "(6) Livelihood", "(7) Capacity")
colnames(at_cor) <- c("(1) Protection", "(2) Management", "(3) Species",
                     "(4) Education", "(5) Law/policy", "(6) Livelihood", "(7) Capacity")

at_cor_plot <- at_cor

# Remove redundant info
at_cor <- at_cor %>%
  format(digits = 2)
at_cor[lower.tri(at_cor)] <- ""
at_cor %>%
  kable()

# Save latex table
kable(at_cor, format = "latex") %>%
  save_kable("actiontypes_cor.tex")

# Build plot
at_cor_plot[lower.tri(at_cor)] <- NA 

at_cor_plot %>% 
  melt(na.rm = TRUE) %>% 
  ggplot(aes(Var2, Var1, fill = value)) +
  geom_tile(color = "white")+
 scale_fill_gradient2(low = "blue", high = "red", mid = "white", 
   midpoint = 0, limit = c(-1,1), space = "Lab", 
   name="Pearson\nCorrelation") +
  theme_minimal()+ 
 theme(axis.text.x = element_text(angle = 45, vjust = 1, 
    size = 12, hjust = 1)) +
  xlab("") + 
  ylab("") +
 coord_fixed()
```

## Additional info

### Strong focus on protection

```{r echo=FALSE, message=FALSE, warning=FALSE}
at_reduced %>% 
  filter(class1 == 2) %>% 
  select(bmz) %>% 
  pull()
  
```

### Strong focus on management

```{r echo=FALSE, message=FALSE, warning=FALSE}
at_reduced %>% 
  filter(class2 == 2) %>% 
  select(bmz) %>% 
  pull()
  
```

### Projects with only one action type

```{r echo=FALSE, message=FALSE, warning=FALSE}
at_reduced_long %>% 
  mutate(max = recode(max,
                      `2` = 1,
                      `1` = 1,
                      `0` = 0)) %>% 
  group_by(bmz) %>% 
  summarize(offocus = sum(max, na.rm = TRUE)) %>% 
  filter(offocus == 1)
  
```

```{r echo=FALSE, message=FALSE, warning=FALSE}
at_reduced_long %>% 
  mutate(max = recode(max,
                      `2` = 1,
                      `1` = 0,
                      `0` = 0)) %>% 
  group_by(bmz) %>% 
  summarize(offocus = sum(max, na.rm = TRUE)) %>% 
  filter(offocus == 1)
  
```

There are no projects with only one action type.

### Projects here vs. projects in draft table

```{r echo=FALSE, message=FALSE, warning=FALSE}
draft_tbl_bmz <- as.character(c(199865809, 199867003, 199867219, 200065086, 200066340, 200165092, 200165472, 200166090, 200266908, 200466094, 200466458, 200565689, 200666131, 200766667, 200865444, 200866285, 200866657, 201065143, 201066836, 201165984, 201166156, 201167063, 201265685, 201367127, 201468594, 201567957, 201670124, 201697515, 209810094, 209810854, 209810961))

at_bmz <- at_reduced %>% 
  select(bmz) %>% 
  pull()

tibble(draft_tbl_bmz) %>%
  filter(draft_tbl_bmz %in% at_bmz)


  
```

There are some projects in the IUCN action type table that are not in Table 1 in the draft paper and vice versa. This should be looked into. 

## How many BMZ per PA

```{r}
count_treatments <- actiontypes %>%  
  group_by(id) %>%  
  summarize(number_bmz = n())
table(count_treatments$number_bmz)
```


```{r echo=FALSE, message=FALSE, warning=FALSE}
at_bybmzpa <- actiontypes %>% 
  unite(col = "bmz_pa", c("bmz", "id"), sep = "_", na.rm = TRUE) %>%
  pivot_longer(c(-bmz_pa)) %>% 
  mutate(name = str_extract(name, "class[0-9]")) %>% 
  group_by(bmz_pa, name) %>% 
  summarise(max = max(value)) %>% 
  # group_by(max, name) %>% 
  # summarise(count = n()) %>% 
  pivot_wider(names_from = "name", values_fill = 0, 
              values_from = "max") %>%
  separate(col = "bmz_pa", into = c("bmz", "id"), sep = "_")

saveRDS(at_bybmzpa, "actiontypes_bybmzpa.RDS")
```

