---
title: "Map - Cells with multiple WDPA IDs"
author: "Yota"
date: "23 6 2022"
output: workflowr::wflow_html
---

```{r setup, include=FALSE, message = FALSE, warning = FALSE}
knitr::opts_chunk$set(echo = FALSE, message = FALSE, warning = FALSE)
knitr::opts_knit$set(root.dir = "~/shared/datalake/mapme.protectedareas")
```

```{r options and workspace, include=FALSE}
# clean workspace, set options
rm(list=ls())
options(scipen=999)

# get packages
lop <- c("tidyverse", "sf", "leaflet", "leaflet.extras", "leaflet.extras2", "ggsci", "scales", "htmltools", "RColorBrewer", "raster", "rgdal", "stars", "kableExtra", "plotly")
newp <- lop[!(lop %in% installed.packages()[,"Package"])]
if(length(newp)) install.packages(newp)
lapply(lop, require, character.only = TRUE)

```


```{r load data, include=FALSE}

##  Protected areas
wdpa_kfw<-
  read_sf("~/shared/datalake/mapme.protectedareas/input/wdpa_kfw/wdpa_kfw_spatial_latinamerica_2021-02-01_supportedPAs_unique.gpkg")

# Load KfW finance data 
kfw_finance <- 
  read_csv("~/shared/datalake/mapme.protectedareas/input/kfw_finance/mapme.protectedareas_kfw-finance-2021-03-17.csv") %>% 
  filter(! is.na(bmz_nummer))

#Load wdpa bmz keys
keys_wdpaid_bmz <- read_csv("~/shared/datalake/mapme.protectedareas/output/matching/model_frames/keys_wdpaid_bmz.csv") %>% 
  rename("bmz_nummer" = "value")

# merge wdpa_kfw with keys
wdpa_kfw <- left_join(wdpa_kfw, keys_wdpaid_bmz,
                      by=c("WDPAID"))

#add kfw_finance data to wdpa_kfw
wdpa_kfw <- left_join(wdpa_kfw, kfw_finance, 
                      by=c("bmz_nummer")) 

wdpa_year <- wdpa_kfw %>% 
  dplyr::select(WDPAID, first_year) %>% 
  st_drop_geometry

# Load fishnet
# fishnet_all <- read_sf("~/shared/datalake/mapme.protectedareas/processing/fishnet/honeycomb_5_sqkm.gpkg")
fishnet_all <- read_sf("~/shared/datalake/mapme.protectedareas/processing/fishnet/honeycomb_5_sqkm_subset_intersect.gpkg") #%>% 
  # select(geom)

fishnet_named <- cbind(rownames(fishnet_all),fishnet_all)

colnames(fishnet_named) <- c(".assetid", colnames(fishnet_all))

# Load link between assetid and WDPA ID
keys_assetid_wdpa <- 
  read_sf("~/shared/datalake/mapme.protectedareas/processing/fishnet/honeycomb_5_sqkm_kfw.gpkg") %>% 
  dplyr::select(poly_id, WDPAID) %>% 
  rename(".assetid" = "poly_id") %>% 
  st_drop_geometry() %>% 
  distinct(.assetid, WDPAID)   %>% 
  mutate(.assetid = as.factor(.assetid))

fishnet_dropgeom <- st_drop_geometry(fishnet_named)

keys_filtered <- keys_assetid_wdpa %>% 
  group_by(.assetid) %>% 
  count() %>% 
  filter(n > 1)

# Merged data for table

check_data_merged <- left_join(wdpa_year, keys_assetid_wdpa,
                               by=c("WDPAID")) %>% 
  filter(first_year > 2003) %>% 
  unite("key", c("WDPAID", ".assetid"), remove = FALSE)

# Merged data for map 

check_data <- 
  left_join(fishnet_dropgeom, keys_assetid_wdpa, 
            by=c(".assetid")) %>% 
  filter(! is.na(WDPAID),
         .assetid %in% keys_filtered$.assetid) %>%  
  left_join(., wdpa_year,
            by=c("WDPAID")) %>% 
  dplyr::select(.assetid, WDPAID, first_year) %>% 
  mutate(key = paste(.assetid, first_year)) %>%  
  left_join(., fishnet_named, 
            by=c(".assetid"))

count <- check_data %>% 
  group_by(key) %>% 
  count()

check_okay <- left_join(check_data, count,
                        by="key") %>% 
  filter(n == 1)

check_data <- left_join(check_data, count,
                        by="key") %>% 
  filter(n > 1)

check_data <- st_sf(check_data) %>% 
  st_transform(., ,crs = st_crs(wdpa_kfw))

check_data_exp <- check_data %>% 
  filter(first_year == 2007)


```

## Problem Description

Reasons for this can be: 

* Cells at the border of multiple PAs
* Cells within overlapping PAs

Only cells that cause issues in our analysis, i.e. cells for which assigned PAs have had a project starting in the same year. 

Example for illustration: Say, we have the following cells:

* Cell "12345" with corresponding PAs with WDPA IDs "1" and "2", and
* Cell "23456" with corresponding PAs with WDPA IDs "3" and "4".

Now, say PAs "1" and "2" both have a project starting in the same year; PAs "3" and "4" have projects starting in different years. In that case, cell "12345" is shown in the map below, cell "23456" is not. Reason: The correct (for our analysis) WDPA ID can be identified for cell "23456" by filtering by project start year. For cell "12345", this is not possible. 

## Table

This table shows the number of *treatment* cells with multiple assigned PAs (WDPA IDs) for each matching frame (year).
 
```{r table}
# Load obs table
observations_table <- readRDS("/datadrive/datalake/mapme.protectedareas/output/matching/output_tables/observations_table.rds")
obs_treated <- as.numeric(observations_table[2,c(2:10, 12, 13, 15)])

# Data frame with variables of interest

collapsed_data <- 
  check_data_merged %>% 
  # filter(first_year == year) %>%
  distinct(key, .keep_all=TRUE) %>% 
  group_by(first_year,.assetid) %>% 
  count() %>% 
  filter(n>1,
         ! is.na(.assetid))

# Create a table 
year_table <- data.frame(table(collapsed_data$first_year),
              obs_treated)
colnames(year_table) <- c("Matching frame (Year)", "Cells with multiple WDPA IDs", "Total number of treated cells")
year_table %>% 
  kbl() %>% 
  kable_classic(full_width = T, html_font = "Cambria")
```

The following plot displays the number of observations vs. the number of WDPA IDs assigned to a cell. This plot does not show cells that have only one assigned WDPA ID. Most cells that have multiple WDPA IDs have only *two* corresponding WDPA IDs. 

```{r}
# Create bar plot

obs_per_bin <- 
  collapsed_data %>% 
  ggplot(aes(x=n, fill=as.factor(first_year), group=as.factor(first_year))) +
  geom_bar(position = "stack") +
  labs(x="Number of WDPA IDs assigned", 
       y="Number of observations",
       fill="Year")
ggplotly(obs_per_bin)

```


## Map 

This map shows *treatment* cells that have multiple assigned PAs (WDPA IDs). 

```{r map}
#------- Color palettes --------



# color palette, which is interpolated ?
# pal <- colorNumeric(palette = "Spectral", #c("#000000", "#666666", "#FFFFFF"), 
#                     values(accessibility),
#                     na.color = "transparent")

# pal <- colorQuantile(palette = "Oranges", #c("#000000", "#666666", "#FFFFFF"),
#                     values(accessibility),
#                     n = 8,
#                     probs = c(seq(0, 0.8, 0.1),1),#seq(0, 1, length.out = 9),
#                     na.color = "transparent")

# # the extra code building off the existing pal
# pal_colors <- unique(pal(sort(values(accessibility)))) # hex codes
# pal_labs <- quantile(values(accessibility), c(seq(0, 0.8, 0.1),1), na.rm = TRUE) # depends on n from pal
# # pal_labs <- paste(lag(pal_labs), pal_labs, sep = " - ")[-1] # first lag is NA

# lab_lower <- pal_labs[1:9]
# lab_upper <- pal_labs[-1]
# pal_labs <- paste(lab_lower, lab_upper, sep = " - ")

# pal <- colorBin(palette = "RdGy", #c("#000000", "#666666", "#FFFFFF"), 
#                     values(accessibility),
#                     bins = 10,
#                     pretty = TRUE,
#                     na.color = "transparent")

#------- Map --------

## Create map
my_map_matching <-
  leaflet() %>%
  # add external map providers
  addTiles(group = "OpenStreetMap") %>%
  addProviderTiles(providers$CartoDB.Positron, group="CartoDB") %>%
  addProviderTiles(providers$Esri.WorldImagery, group="Satellite") %>%
  addProviderTiles(providers$Esri.WorldShadedRelief, group="Topography") %>%
  addProviderTiles(providers$NASAGIBS.ViirsEarthAtNight2012, group="Nightlights") %>%
  # add own data
  addPolygons(data = wdpa_kfw,
              opacity = 0.8,
              color = "orange", 
              group = "PA Boundaries",
              label = ~htmlEscape(DESIG_ENG),
              weight = 1) %>%
  addPolygons(data = check_data,
              opacity = 0.8,
              color = "green", 
              group = "Problematic cells",
              label = ~htmlEscape(.assetid),
              weight = 1) %>%
  addPolygons(data = check_data_exp,
              opacity = 0.8,
              color = "green", 
              group = "Problematic cells (Example year)",
              label = ~htmlEscape(.assetid),
              weight = 1) %>%
  addFullscreenControl() %>%
  # add layers control to define which data is shown or ommited in default view
  addLayersControl(
    baseGroups = c("CartoDB","OpenStreetMap","Satellite","Topography","Nightlights"), #"Toner",,"Regional Primary Forests (2001)"
    overlayGroups = c("PA Boundaries", "Problematic cells", "Problematic cells (Example year)"),
    options = layersControlOptions(collapsed = FALSE)) #%>%
    # omit certain layers
    # hideGroup(group = c())

my_map_matching


```