---
title: "Map of protected areas with financial support from KfW"
author: "Johannes Schielein (Author)"
date: "2021-12-22"
output: workflowr::wflow_html
editor_options:
  chunk_output_type: console
---

```{r setup, message = FALSE,warning = FALSE,include = FALSE}
# load relevant libraries
library("tidyverse")
library("sf")
library("leaflet")
library("leaflet.extras")
library("leaflet.extras2")
library("ggsci")
library("scales")
library("htmltools")
library("RColorBrewer")
```

```{r database_creation, message = FALSE, warning = FALSE, include = FALSE}
##  Protected areas
##  Protected areas
wdpa_kfw<-
  read_sf("~/shared/datalake/mapme.protectedareas/input/wdpa_kfw/wdpa_kfw_spatial_latinamerica_2021-02-01_supportedPAs_unique.gpkg")

## create column for area coloring
wdpa_kfw$REP_AREA_cat<-
  cut(wdpa_kfw$REP_AREA,
      c(0,1000,5000,10000,20000,max(wdpa_kfw$REP_AREA)),
      c("< 1,000 sqkm","1,001-5,000 sqkm","5,001-10,000 sqkm","10,001-20,000 sqkm",paste("20,001-",max(wdpa_kfw$REP_AREA)," sqkm",sep="")))


## fishnet
fishnet <-
  read_sf(
    "../../datalake/mapme.protectedareas/output/polygon/sampling/fishnets/fishnet_all_update_Dec-07.gpkg"
  )
# load matching frame data
# matched_data<-rbind(read.csv("../../datalake/mapme.protectedareas/output/tabular/regression_input/matched_panel_2015.csv"),
#                     read.csv("../../datalake/mapme.protectedareas/output/tabular/regression_input/matched_panel_2007.csv"))

matched_data<-read.csv("../../datalake/mapme.protectedareas/output/tabular/regression_input/matched_panel_2015.csv")
# filter to delete multiple observations
matched_data<-
  matched_data %>%
  filter(year == 2015)
# merge data 
matched_data_merged<-
  merge(fishnet, matched_data, "poly_id")

matched_data_merged<-st_transform(matched_data_merged,crs = st_crs(wdpa_kfw))

## Create Color Pals for the plot data
# create colorramp function for area
pal_area <- colorFactor(
  palette = pal_npg("nrc", alpha = 0.7)(length(unique(wdpa_kfw$REP_AREA_cat))),
  domain = wdpa_kfw$REP_AREA_cat
)

# create colorramp2
pal_treatment <- colorFactor(
  palette = c("darkblue","orange"),
  domain = matched_data_merged$treat_ever
)

# create colorramp function for country
colourCount = length(unique(wdpa_kfw$ISO3))
getPalette = colorRampPalette(brewer.pal(9, "Set1"))

pal_country <- colorFactor(
  palette = getPalette(colourCount),
  domain = wdpa_kfw$ISO3
)

```

## Map of Supported Areas

We machted our Latin America portfolio with the *World Database on Protected Areas - WDPA* (IUCN) and used data from the multiple different open data-sources to make an assessment of our portfolio and evaluate the impacts of our projects. Our database currently comprises `r nrow(wdpa_kfw)` PAs in Latin America which are situated in `r length(unique(wdpa_kfw$ISO3))` different countries. Those areas can be broadly categorized into `r table(wdpa_kfw$MARINE)[[3]]` terrestrial, `r table(wdpa_kfw$MARINE)[[1]]` marine, and `r table(wdpa_kfw$MARINE)[[2]]` partial marine/terrestrial protected areas. The georeferenced data covers a total surface of `r round(sum(st_area(wdpa_kfw))/(10^12),digits=3)` Mio. km^2^ which is about `r round(sum(st_area(wdpa_kfw))/(10^9)/357.581,digits=1)` times the size of Germany. 

```{r projectmap, echo=FALSE, warning=FALSE,  fig.width=8}
my_map <-
  leaflet() %>%
  # add external map providers
  addTiles(group = "OpenStreetMap") %>%
  addProviderTiles(providers$CartoDB.Positron, group="CartoDB") %>%
  addProviderTiles(providers$Esri.WorldImagery, group="Satellite") %>%
  addProviderTiles(providers$Esri.WorldShadedRelief, group="Topography") %>%
  addProviderTiles(providers$NASAGIBS.ViirsEarthAtNight2012, group="Nightlights") %>%
  addTiles("https://tiles.globalforestwatch.org/umd_tree_cover_loss/latest/dynamic/{z}/{x}/{y}.png",group="Forest Cover Loss (2001-2020)",attribution = "Hansen, M. C., P. V. Potapov, R. Moore, M. Hancher, S. A. Turubanova, A. Tyukavina, D. Thau, S. V. Stehman, S. J. Goetz, T. R. Loveland, A. Kommareddy, A. Egorov, L. Chini, C. O. Justice, and J. R. G. Townshend. 2013. “High-Resolution Global Maps of 21st-Century Forest Cover Change.” Science 342 (15 November): 850–53. Data available on-line from: http://earthenginepartners.appspot.com/science-2013-global-forest.")%>%
#  addTiles("https://tiles.globalforestwatch.org/umd_regional_primary_forest_2001/latest/dynamic/{z}/{x}/{y}.png",group="Regional Primary Forests (2001)",attribution = "Hansen, M. C., P. V. Potapov, R. Moore, M. Hancher, S. A. Turubanova, A. Tyukavina, D. Thau, S. V. Stehman, S. J. Goetz, T. R. Loveland, A. Kommareddy, A. Egorov, L. Chini, C. O. Justice, and J. R. G. Townshend. 2013. “High-Resolution Global Maps of 21st-Century Forest Cover Change.” Science 342 (15 November): 850–53. Data available on-line from: http://earthenginepartners.appspot.com/science-2013-global-forest.")%>%
  # add own data
  addPolygons(data = wdpa_kfw,opacity = 0.9,color = "orange", group = "PA Boundaries (all years)",label = ~htmlEscape(NAME),weight = 1)%>%
  addPolygons(data = wdpa_kfw,opacity = 1,color = ~pal_area(REP_AREA_cat), group = "PA Area Size",label = ~htmlEscape(REP_AREA),weight = 1)%>%
  addPolygons(data = wdpa_kfw,opacity = 1,color = ~pal_country(ISO3), group = "Country",label = ~htmlEscape(ISO3),weight = 1)%>%
  addPolygons(data = matched_data_merged, opacity = 0.9,color = ~pal_treatment(treat_ever), group = "Cells (Treamtent & Control) in 2015",label = ~htmlEscape(treat_ever),weight = 1)%>%
  # fullscreen control
  addFullscreenControl() %>%
  # add legent for area
  addLegend("bottomright",
            data = wdpa_kfw,
            pal = pal_area,
            values = ~REP_AREA_cat,
            title = "Total Reported Area",
            opacity = 1,
            group = "PA Area Size") %>% 
  addLegend("bottomright",
            data = matched_data_merged,
            pal = pal_treatment,
            values = ~treat_ever,
            title = "Treatement",
            opacity = 1,
            group = "Cells (Treamtent & Control) in 2015") %>% 
  addLegend("bottomright",
            data = wdpa_kfw,
            pal = pal_country,
            values = ~ISO3,
            title = "Country",
            opacity = 1,
            group = "Country") %>% 
  # add layers control to define which data is shown or ommited in default view
  addLayersControl(
    baseGroups = c("CartoDB","OpenStreetMap","Satellite","Topography","Nightlights","Forest Cover Loss (2001-2020)"), #"Toner","Regional Primary Forests (2001)"
    overlayGroups = c("PA Boundaries (all years)","Country","PA Area Size","Cells (Treamtent & Control) in 2015"),
    options = layersControlOptions(collapsed = FALSE)) %>%
  # ommit certain layers
  hideGroup(group = c("Country","PA Area Size","Cells (Treamtent & Control) in 2015"))

my_map
```