---
title: "Map of protected areas with technical support from GIZ"
author: "Johannes Schielein (Author)"
date: "2022-02-11"
output: workflowr::wflow_html
editor_options:
  chunk_output_type: console
---

```{r setup, message = FALSE,warning = FALSE,include = FALSE}
# load relevant libraries
library("tidyverse")
library("sf")
library("leaflet")
library("leaflet.extras")
library("leaflet.extras2")
library("ggsci")
library("scales")
library("htmltools")
library("RColorBrewer")
```

```{r database_creation, message = FALSE, warning = FALSE, include = FALSE}
##  Protected areas
##  Protected areas
wdpa_giz<-
  read_sf("~/shared/datalake/mapme.protectedareas/input/wdpa_giz/wdpa_giz_supported_2019_wdpaV_Feb2022.gpkg")

wdpa_kfw<-
  read_sf("~/shared/datalake/mapme.protectedareas/input/wdpa_kfw/wdpa_kfw_spatial_latinamerica_2021-02-01_supportedPAs_unique.gpkg")


## create column for area coloring
wdpa_giz$REP_AREA_cat<-
  cut(wdpa_giz$REP_AREA,
      c(0,1000,5000,10000,20000,max(wdpa_giz$REP_AREA)),
      c("< 1,000 sqkm","1,001-5,000 sqkm","5,001-10,000 sqkm","10,001-20,000 sqkm",paste("20,001-",max(wdpa_giz$REP_AREA)," sqkm",sep="")))

## Create Color Pals for the plot data
# create colorramp function for area
pal_area <- colorFactor(
  palette = pal_npg("nrc", alpha = 0.7)(length(unique(wdpa_giz$REP_AREA_cat))),
  domain = wdpa_giz$REP_AREA_cat
)

# create colorramp2
# pal_treatment <- colorFactor(
#   palette = c("darkblue","orange"),
#   domain = matched_data_merged$treat_ever
# )

# create colorramp function for country
colourCount = length(unique(wdpa_giz$ISO3))
getPalette = colorRampPalette(brewer.pal(9, "Set1"))

pal_country <- colorFactor(
  palette = getPalette(colourCount),
  domain = wdpa_giz$ISO3
)

```

## Map of Supported Areas

The following map shows a match of the GIZ portfolio with the *World Database on Protected Areas - WDPA*. [^1]. The displayed database currently comprises `r nrow(wdpa_giz)` PAs which are situated in `r length(unique(wdpa_giz$ISO3))` different countries. Those areas can be broadly categorized into `r table(wdpa_giz$MARINE)[[1]]` terrestrial, `r table(wdpa_giz$MARINE)[[2]]` marine, and `r table(wdpa_giz$MARINE)[[3]]` partly marine/terrestrial protected areas. The georeferenced data covers a total surface of `r round(sum(st_area(wdpa_giz))/(10^12),digits=3)` Mio. km^2^ which corresponds to about `r round(sum(st_area(wdpa_giz))/(10^9)/357.581,digits=1)` times the size of Germany. 

Feel free to play around with the map and activate different layers. 
```{r projectmap, echo=FALSE, warning=FALSE,  fig.width=8}
## Crate map
# wdpa_giz_treatment_centroid<-st_transform(st_centroid(wdpa_giz_treatment_centroid),crs = 4326)
my_map <-
  leaflet() %>%
  # add external map providers
  addTiles(group = "OpenStreetMap") %>%
  addProviderTiles(providers$CartoDB.Positron, group="CartoDB") %>%
  addProviderTiles(providers$Esri.WorldImagery, group="Satellite") %>%
  addProviderTiles(providers$Esri.WorldShadedRelief, group="Topography") %>%
  addProviderTiles(providers$NASAGIBS.ViirsEarthAtNight2012, group="Nightlights") %>%
  addTiles("https://tiles.globalforestwatch.org/umd_tree_cover_loss/latest/dynamic/{z}/{x}/{y}.png",
           group="Forest Cover Loss (2001-2020)",
           #options=tileOptions(opacity = 0.7),
           attribution = "Hansen, M. C., P. V. Potapov, R. Moore, M. Hancher, S. A. Turubanova, A. Tyukavina, D. Thau, S. V. Stehman, S. J. Goetz, T. R. Loveland, A. Kommareddy, A. Egorov, L. Chini, C. O. Justice, and J. R. G. Townshend. 2013. “High-Resolution Global Maps of 21st-Century Forest Cover Change.” Science 342 (15 November): 850–53. Data available on-line from: http://earthenginepartners.appspot.com/science-2013-global-forest.")%>%
  # add own data
    addPolygons(data = wdpa_giz,opacity = 0.9,color = "red", group = "GIZ Supported PAs",label = ~htmlEscape(NAME),weight = 1)%>%
      addPolygons(data = wdpa_kfw,opacity = 0.9,color = "#2b8cbe", group = "KfW Supported PAs (Latinamerica)",label = ~htmlEscape(NAME),weight = 1)%>%
  addPolygons(data = wdpa_giz,opacity = 1,color = ~pal_area(REP_AREA_cat), group = "PAs by Area Size",label = ~htmlEscape(REP_AREA),weight = 1)%>%
  addPolygons(data = wdpa_giz,opacity = 1,color = ~pal_country(ISO3), group = "PAs by Country",label = ~htmlEscape(ISO3),weight = 1)%>%
  # fullscreen control
  addFullscreenControl() %>%
  # add legent for area
  addLegend("bottomright",
            data = wdpa_giz,
            pal = pal_area,
            values = ~REP_AREA_cat,
            title = "Total Reported Area",
            opacity = 1,
            group = "PAs by Area Size") %>% 
  addLegend("bottomright",
            data = wdpa_giz,
            pal = pal_country,
            values = ~ISO3,
            title = "Country",
            opacity = 1,
            group = "PAs by Country") %>% 
  # add layers control to define which data is shown or ommited in default view
  addLayersControl(
    baseGroups = c("CartoDB","OpenStreetMap","Satellite","Topography","Nightlights"), #"Toner",,"Regional Primary Forests (2001)"
    overlayGroups = c("GIZ Supported PAs","PAs by Country","PAs by Area Size","Forest Cover Loss (2001-2020)","KfW Supported PAs (Latinamerica)"),
    options = layersControlOptions(collapsed = FALSE)) %>%
  # ommit certain layers
  hideGroup(group = c("PAs by Country","PAs by Area Size","Forest Cover Loss (2001-2020)","KfW Supported PAs (Latinamerica)"))

my_map
```


[comment]: <> (Literature)

[^1]: UNEP-WCMC and IUCN (2022), Protected Planet: The World Database on Protected Areas (WDPA) and World Database on Other Effective Area-based Conservation Measures (WD-OECM) [Online], February 2022, Cambridge, UK: UNEP-WCMC and IUCN. Available at: www.protectedplanet.net.
