---
title: "Threat Assessment for Protected Areas"
author: "Johannes Schielein (Author)"
date: "2022-02-10"
output: workflowr::wflow_html
editor_options:
  chunk_output_type: console
---

```{r setup, message = FALSE, warning = FALSE, include = FALSE}
# load libraries and set session options. 
library("tidyverse")
library("sf")
library("leaflet")
library("leaflet.extras")
library("leaflet.extras2")
library("ggsci")
library("scales")
library("htmltools")
library("RColorBrewer")
library("plotly")
options(scipen=10000)
```

```{r database_creation, message = FALSE, warning = FALSE, include = FALSE}
# ----- load and transform protected areas data -----
##  Protected areas
wdpa_kfw<-
  read_sf("~/shared/datalake/mapme.protectedareas/input/wdpa_kfw/wdpa_kfw_spatial_latinamerica_2021-02-01_supportedPAs_unique_simplified.gpkg")

## create column for area coloring based on categories
wdpa_kfw$REP_AREA_cat<-
  cut(wdpa_kfw$REP_AREA,
      c(0,1000,5000,10000,20000,max(wdpa_kfw$REP_AREA)),
      c("< 1,000 sqkm","1,001-5,000 sqkm","5,001-10,000 sqkm","10,001-20,000 sqkm",paste("20,001-",max(wdpa_kfw$REP_AREA)," sqkm",sep="")))

# ----- load loss data and prepare for mapping -----
## loss statistics
gfw_lossstats <-
  read_csv(
    "../../datalake/mapme.protectedareas/output/polygon/global_forest_watch/zonal_statistics_allPAs_long_temporal.csv",
    col_types = c("cfid")
  )

## filter for area only
gfw_lossstats <-
  gfw_lossstats %>%
  filter(name == "area")

## load spatial data
wdpa_allPAs<-
  st_read("../../datalake/mapme.protectedareas/input/wdpa_kfw/wdpa_kfw_spatial_latinamerica_2021-04-22_allPAs_valid_simplified.gpkg")

## check if wdpa ids from loss statistics are contained in spatial data
table(unique(gfw_lossstats$WDPA_PID)%in%wdpa_allPAs$WDPA_PID) # all but one

## add column to spatial data that shows forest area in 2000 for each area (maximum is equal to area in 2000 since no forest gain is in the data)
max_forest <-
  gfw_lossstats %>%
  group_by(WDPA_PID) %>%
  dplyr::summarise(max_forest = max(value))

## merge to spatial and clean up
wdpa_allPAs <- 
  merge(wdpa_allPAs, max_forest,"WDPA_PID")

rm(max_forest)

## categorize maximum forest area by meaningfull and pretty breaks
wdpa_allPAs$max_forest_categorized <-
  cut(
    wdpa_allPAs$max_forest,
    breaks = c(-1, 100, 1000, 10000,100000, max(wdpa_allPAs$max_forest, na.rm = T)),
    labels = c("0-100", "101-1,000", "1,001-10,000", "10,001-100,000",">100,000")
  )

## calculate relative losses between 2000 and 2020 and add column to spatial data
relative_loss <-
  gfw_lossstats %>%
  group_by(WDPA_PID) %>%
  dplyr::summarise(relative_loss = 1-min(value)/max(value),absolute_loss = max(value)-min(value))

## merge to spatial and clean up
wdpa_allPAs <- 
  merge(wdpa_allPAs, relative_loss, "WDPA_PID")

rm(relative_loss)

## categorize relative loss by meaningfull and pretty breaks 
wdpa_allPAs$relative_loss_categorized <-
  cut(
    wdpa_allPAs$relative_loss,
    breaks = c(-1, 0.02, 0.05, 0.1, 0.2, 1),
    labels = c("0-2 %", "2-5 %", "5-10 %", "10-20 %",">20 %")
  )


## categorize absolute loss by meaningfull and pretty breaks 
wdpa_allPAs$absolute_loss_categorized <-
  cut(
    wdpa_allPAs$absolute_loss,
    breaks = c(-1, 10, 100, 1000, 10000,max(wdpa_allPAs$absolute_loss)),
    labels = c("0-10", "10-100", "100-1,000", "1,000-10,000",">10,000")
  )

## subset data and get centroids for map
wdpa_allPAs_lossdata<-wdpa_allPAs %>% 
  filter(DESIG_ENG != 'UNESCO-MAB Biosphere Reserve') %>% 
  filter(GEOMETRY_TYPE != 'POINT') %>% 
  filter(is.nan(relative_loss)!=TRUE) %>%
  st_centroid()

## subset all data for untreated
wdpa_nonSupportPA<-wdpa_allPAs %>% 
  filter(DESIG_ENG != 'UNESCO-MAB Biosphere Reserve') %>% 
  filter(GEOMETRY_TYPE != 'POINT') %>% 
  filter(is.na(bmz_n_1)==TRUE) 

```
## Introduction
The following graphics are intented to support a peer-discussion on how publicly available geodata can be used to characterize the threat level of protected areas (short PAs). The goal of this exercise is to define data products that could support the identification of new intervention sites and thus help KfW and its partners in the project preparation phase. In addition, information about the current portfolio and its relevance to protect the most threatened areas can be derived.

The general idea of this analysis is to match the complete network of existing PAs in countries where KfW is active to the the database of supported PAs by KfW. The same methodology can be used later in the project preparation phase to analyse the existing portfolio together with newly projected intervention sites. The analysis is based on publicly available data from the World Database of Protected Areas (WDPA/IUCN).[^1] Our current portfolio database contains `r nrow(wdpa_allPAs[!is.na(wdpa_allPAs$bmz_n_1),])` areas from a total of `r nrow(wdpa_allPAs)` in the partnering countries. 

In a first analysis step we focus on habitat destruction in terms of primary forest cover loss. To quantify forest cover loss we utilize data from the Global Forest Watch (Hansen et al, 2013)[^2]. Forest cover loss is defined in their work _"as a stand-replacement disturbance, or a change from a forest to non-forest state."_. Loss can either be the result of human activities or natural factors such as droughts, forest fires or hurricanes, amongst others. More information on the interpretation and usefullness of this data as well as suggested further steps to advance the threat assessment are given below in the discussion part. 

__In the analysis below we will focus on two key outcome indicators:__ 

* **Total forest cover loss**: Measures the total sum of loss areas in hectare. This variable is able to identify PAs with the highest primary forest cover loss between 2001 and 2020. The identification of high loss areas can be usefull for targeting areas where we might achieve the largest impact in terms of _reducing emissions from deforstation and forest degradation_. 

* **Relative forest cover loss**: Measures the percentage of primary forest cover loss inside a PA compared to its total primary forest area in 2000. The identification of PAs with high relative losses can be relevant from a biodiversity perspective. PAs with high relative losses might be places where large parts of the functional forest habitat is lost. Targeting these areas might not only help to protect the _floral biodiversity but also the fauna and humans that inhabit these areas and profit from the local forest ecosystem services_. 

## Map of Threatened Areas

```{r projectmap, echo=FALSE, warning=FALSE,  fig.width=10}
## ----- create color palettes for the map -----
pal_relative_loss <- colorFactor(
  palette = rev(brewer.pal(5, "RdYlGn")),
  domain = wdpa_allPAs$relative_loss_categorized
)

pal_absoute_loss <- colorFactor(
  palette = rev(brewer.pal(5, "RdYlGn")),
  domain = wdpa_allPAs$absolute_loss_categorized
)

# create colorramp function for country
colourCount = length(unique(wdpa_kfw$ISO3))
getPalette = colorRampPalette(brewer.pal(9, "Set1"))

pal_country <- colorFactor(
  palette = getPalette(colourCount),
  domain = wdpa_kfw$ISO3
)

# create colorramp2
# pal_treatment <- colorFactor(
#   palette = c("darkblue","orange"),
#   domain = matched_data_merged$treat_ever
# )

## add label string for map display 
wdpa_allPAs_lossdata$label <- 
  with(wdpa_allPAs_lossdata, paste(
  "<p> <b>", NAME, "</b> </br>",
  "Primary Forests (2000)",round(max_forest,digits=0)," ha", "</br>",
  "Absolute Loss (2000-2020):",round(absolute_loss,digits=0)," ha", "</br>",
  "Relative Loss (2000-2020):", round(relative_loss*100,digits=2)," %:",
  "</p>"))

# ----- create map -----
my_map <-
  leaflet() %>%
  # add external map providers
  addTiles(group = "OpenStreetMap") %>%
  addProviderTiles(providers$CartoDB.Positron, group="CartoDB") %>%
  addProviderTiles(providers$Esri.WorldImagery, group="Satellite") %>%
  addProviderTiles(providers$Esri.WorldShadedRelief, group="Topography") %>%
  addProviderTiles(providers$NASAGIBS.ViirsEarthAtNight2012, group="Nightlights") %>%
  addTiles(
    "https://tiles.globalforestwatch.org/umd_tree_cover_loss/latest/dynamic/{z}/{x}/{y}.png",
    group = "Forest Cover Loss (2001-2020)",
    #options=tileOptions(opacity = 0.7),
    attribution = "Hansen, M. C., P. V. Potapov, R. Moore, M. Hancher, S. A. Turubanova, A. Tyukavina, D. Thau, S. V. Stehman, S. J. Goetz, T. R. Loveland, A. Kommareddy, A. Egorov, L. Chini, C. O. Justice, and J. R. G. Townshend. 2013. “High-Resolution Global Maps of 21st-Century Forest Cover Change.” Science 342 (15 November): 850–53. Data available on-line from: http://earthenginepartners.appspot.com/science-2013-global-forest."
  ) %>%
  addTiles(
    "https://tiles.globalforestwatch.org/umd_regional_primary_forest_2001/latest/dynamic/{z}/{x}/{y}.png",
    group = "Regional Primary Forests (2001)",
    attribution = "Hansen, M. C., P. V. Potapov, R. Moore, M. Hancher, S. A. Turubanova, A. Tyukavina, D. Thau, S. V. Stehman, S. J. Goetz, T. R. Loveland, A. Kommareddy, A. Egorov, L. Chini, C. O. Justice, and J. R. G. Townshend. 2013. “High-Resolution Global Maps of 21st-Century Forest Cover Change.” Science 342 (15 November): 850–53. Data available on-line from: http://earthenginepartners.appspot.com/science-2013-global-forest."
  ) %>%
  # add own polygon data
  addPolygons(data = wdpa_kfw,opacity = 0.9,color = "#2b8cbe", group = "PAs KfW-supported",label = ~htmlEscape(NAME),weight = 1)%>%
  addPolygons(data = wdpa_nonSupportPA,opacity = 0.7,color = "darkgrey", group = "PAs (Others)",label = ~htmlEscape(NAME),weight = 1)%>%
  # add own forest loss data as circles
  addCircleMarkers(data=wdpa_allPAs_lossdata,
                   color = ~pal_relative_loss(relative_loss_categorized),
                   group = "Relative Forest Loss (2001-2020)",
                   radius = ~ifelse(absolute_loss_categorized == ">10,000", 12, 
                                    ifelse(absolute_loss_categorized == "1,000-10,000", 6,
                                           ifelse(absolute_loss_categorized == "100-1,000", 4,
                                                  ifelse(absolute_loss_categorized == "10-100", 2,1
                                                  )      
                                           )
                                    )),
                   popup = ~label,
                   stroke = FALSE, 
                   fillOpacity = 0.5
                   ) %>%
  # add fullscreen control
  addFullscreenControl() %>%
  # add legend(s)
  addLegend("bottomright",
            data = wdpa_allPAs,
            pal = pal_relative_loss,
            values = ~relative_loss_categorized,
            title = "Relative Forest Loss (2001-2020)",
            opacity = 1,
            group = "PA Area Size") %>%
  
  # add layers control to define how and where data is displayed.
  addLayersControl(
    baseGroups = c("CartoDB","OpenStreetMap","Satellite","Topography","Nightlights"), #"Toner",,"Regional Primary Forests (2001)"
    overlayGroups = c("PAs KfW-supported","PAs (Others)","Relative Forest Loss (2001-2020)","Forest Cover Loss (2001-2020)", "Regional Primary Forests (2001)"),
    options = layersControlOptions(collapsed = FALSE)) %>%
  # ommit certain layers
  hideGroup(group = c("PA Area Size","Forest Cover Loss (2001-2020)", "Regional Primary Forests (2001)"))

my_map


```

## Lollipop plot 

```{r lollipop_plot, message = FALSE, warning = FALSE,fig.width=10, echo=FALSE}
# ----- create horizontal lollipop plots for top loss areas -----
lossdata_lollipop_plot <-
  wdpa_allPAs_lossdata %>%
  top_n(30, wdpa_allPAs_lossdata$absolute_loss) %>%
  # arrange(desc(absolute_loss)) %>%
  ggplot() +
  geom_point(aes(
    x = reorder(NAME,-absolute_loss),
    y = absolute_loss,
    color = factor(bmz_n_1),
    text = paste(
      "Area Name: ", NAME,
      "\nForest cover loss: ", round(absolute_loss,digits = 0), " ha",
      "\nBMZ-Number: ", bmz_n_1)
  )) +
  geom_segment(aes(
    x = reorder(NAME,-absolute_loss),
    xend = reorder(NAME,-absolute_loss),
    y = 0,
    yend = absolute_loss,
    color = factor(bmz_n_1),
    text = paste(
      "Area Name: ", NAME,
      "\nForest cover loss: ", round(absolute_loss,digits = 0), " ha",
      "\nBMZ-Number: ", bmz_n_1)
  )) +
  coord_flip() +
  labs(x="", y="Absolute Forest Cover Loss in ha (2000-2020)", color = "Project Number")+
  theme_classic()

ggplotly(lossdata_lollipop_plot,tooltip = "text")


```

```{r filter_data, message = FALSE, warning = FALSE, include=FALSE}
## Crosstables
# filter data for seperate display of jitters
wdpa_allPAs_lossdata$treated<-
  ifelse(is.na(wdpa_allPAs_lossdata$bmz_n_1)==TRUE,0,1)

wdpa_allPAs_lossdata_treated<-
  wdpa_allPAs_lossdata %>%
  filter(treated==1)
```
## Crosstables
```{r cross_tables_loss, message = FALSE, warning = FALSE, echo=FALSE, results='asis'}
# ----- crosstables for relative and absolute loss -----
library("knitr")
# create absolute loss table
loss_data_summarycats_absolute <-
  wdpa_allPAs_lossdata %>%
  filter(treated == 0) %>%
  {
    round(table(.$absolute_loss_categorized) / nrow(.) * 100, digits = 2)
  } %>%
  as.data.frame() %>%
  rename("Absolute Forst Loss (ha)" = Var1, "% of all PAs" = Freq) 

loss_data_summarycats_absolute$`% of KfW Supported PAs` <-
  wdpa_allPAs_lossdata %>%
  filter(treated == 1) %>%
  {
    round(table(.$absolute_loss_categorized) / nrow(.) * 100, digits = 2)
  } %>%
  as.data.frame() %>%
  rename("Absolute Forst Loss (ha)" = Var1,
         "% of KfW Supported PAs" = Freq) %>%
  pull("% of KfW Supported PAs")

kable(loss_data_summarycats_absolute)

# create relative loss table
loss_data_summarycats_relative <-
  wdpa_allPAs_lossdata %>%
  filter(treated == 0) %>%
  {
    round(table(.$relative_loss_categorized) / nrow(.) * 100, digits = 2)
  } %>%
  as.data.frame() %>%
  rename("Relative Forst Loss" = Var1, "% of all PAs" = Freq) 

loss_data_summarycats_relative$`% of KfW Supported PAs` <-
  wdpa_allPAs_lossdata %>%
  filter(treated == 1) %>%
  {
    round(table(.$relative_loss_categorized) / nrow(.) * 100, digits = 2)
  } %>%
  as.data.frame() %>%
  rename("Relative Forst Loss (ha)" = Var1,
         "% of KfW Supported PAs" = Freq) %>%
  pull("% of KfW Supported PAs")

kable(loss_data_summarycats_relative)


```
## Jitter Plots
```{r jitter_plot, message = FALSE, warning = FALSE,fig.width=10, echo=FALSE}
# ----- jitter plots absolute loss -----
# create plot
lossdata_absolute_jitter_plot <-
  wdpa_allPAs_lossdata %>%
  filter(treated == 0) %>%
  ggplot() +
  geom_jitter(
    aes(
    x = ISO3,
    y = absolute_loss,
    color = factor(ISO3),
    text = paste(
      "Area Name: ", NAME,
      "\nAbsolute forest loss: ", round(absolute_loss,digits = 0), " ha",
      "\nRelative forest loss: ", round(relative_loss,digits = 2)*100, " %",
      "\nBMZ-Number: ", bmz_n_1)
    ),
    alpha = .8) +
  geom_jitter(
    data = wdpa_allPAs_lossdata_treated,
    aes(x = ISO3, 
        y = absolute_loss,
        text = paste(
          "Area Name: ", NAME,
          "\nAbsolute forest loss: ", round(absolute_loss,digits = 0), " ha",
          "\nRelative forest loss: ", round(relative_loss,digits = 2)*100, " %",
          "\nBMZ-Number: ", bmz_n_1)),
    color = "black",
    alpha = .8
  )+
  labs(color="Country",x="")+
  scale_y_continuous(labels=function(x) format(x, big.mark = ",", scientific = FALSE),name="Forest Loss in ha")+
  theme_classic()

ggplotly(lossdata_absolute_jitter_plot,tooltip = "text")

# ----- jitter plots relative loss -----
# create plot
lossdata_relative_jitter_plot <-
  wdpa_allPAs_lossdata %>%
  filter(treated == 0) %>%
  ggplot() +
  geom_jitter(
    aes(
      x = ISO3,
      y = relative_loss,
      color = factor(ISO3),
      text = paste(
        "Area Name: ", NAME,
        "\nAbsolute forest loss: ", round(absolute_loss,digits = 0), " ha",
        "\nRelative forest loss: ", round(relative_loss,digits = 2)*100, " %",
        "\nBMZ-Number: ", bmz_n_1)
    ),
    alpha = .8) +
  geom_jitter(
    data = wdpa_allPAs_lossdata_treated,
    aes(x = ISO3, 
        y = relative_loss,
        text = paste(
          "Area Name: ", NAME,
          "\nAbsolute forest loss: ", round(absolute_loss,digits = 0), " ha",
          "\nRelative forest loss: ", round(relative_loss,digits = 2)*100, " %",
          "\nBMZ-Number: ", bmz_n_1)),
    color = "black",
    alpha = .8
  )+
  labs(color="Country",x="")+
  scale_y_continuous(labels=function(x) format(x, big.mark = ",", scientific = FALSE),name="Forest Loss in ha")+
  theme_classic()

ggplotly(lossdata_absolute_jitter_plot,tooltip = "text")

```
## Discussion

[comment]: <> (Literature)

[^1]: UNEP-WCMC and IUCN (2022), Protected Planet: The World Database on Protected Areas (WDPA) and World Database on Other Effective Area-based Conservation Measures (WD-OECM) [Online], February 2022, Cambridge, UK: UNEP-WCMC and IUCN. Available at: www.protectedplanet.net.

[^2]: "Hansen, M. C., P. V. Potapov, R. Moore, M. Hancher, S. A. Turubanova, A. Tyukavina, D. Thau, S. V. Stehman, S. J. Goetz, T. R. Loveland, A. Kommareddy, A. Egorov, L. Chini, C. O. Justice, and J. R. G. Townshend. 2013. “High-Resolution Global Maps of 21st-Century Forest Cover Change.” Science 342 (15 November): 850–53. Data available on-line from: http://earthenginepartners.appspot.com/science-2013-global-forest."
