---
title: "Project Preparation: Threat Assessment for Protected Areas in Bolivia"
author: "Johannes Schielein & Andreas Petutschnig"
date: "2023-01-24"
output: workflowr::wflow_html
editor_options:
  chunk_output_type: console
---

```{r setup, message = FALSE, warning = FALSE, include = FALSE}
# load libraries and set session options. 
library("tidyverse")
library("sf")
library("leaflet")
library("leaflet.extras")
library("leaflet.extras2")
library("ggsci")
library("scales")
library("htmltools")
library("RColorBrewer")
library("plotly")
options(scipen=10000)
```

```{r database_creation, message = FALSE, warning = FALSE, include = FALSE}
# case specific IDs
bol_wdpa_ids <- c(303891, 555592639, 555592671, 555592663, 555592664, 342481, 9308,
              342482, 342484, 555592677, 555592678, 555592685, 342466,
              555592641, 555592656, 555592600, 342494, 342495, 342496, 342497,
              342498, 555592646, 555592673, 555592608, 20037, 31, 2037, 303887,
              9779, 303884, 303894, 98183, 303885, 303883, 30, 342464, 35,
              342483, 555592593, 342484, 555592669, 342490, 555592668, 342465,
              555592608)

# ----- load and transform protected areas data -----
##  Protected areas
wdpa_kfw<-
  read_sf("~/shared/datalake/mapme.protectedareas/input/wdpa_kfw/wdpa_kfw_spatial_latinamerica_2021-04-22_allPAs_valid_simplified.gpkg")

wdpa_kfw <- wdpa_kfw %>%
  filter(ISO3 == "BOL" &
         DESIG_ENG != "UNESCO-MAB Biosphere Reserve" &
         STATUS != "Proposed"
        )

# Those Regions are only points in the original data and show up as buffered circles in the map
wdpa_noarea <- c(555558388, 555558386, 555558387, 166865, 900565, 900566, 900567, 900783)
wdpa_kfw <- wdpa_kfw %>%
  filter(!WDPAID %in% wdpa_noarea &
         WDPAID %in% bol_wdpa_ids)

## create column for area coloring based on categories
wdpa_kfw$REP_AREA_cat<-
  cut(wdpa_kfw$REP_AREA,
      c(0,1000,5000,10000,20000,max(wdpa_kfw$REP_AREA)),
      c("< 1,000 sqkm","1,001-5,000 sqkm","5,001-10,000 sqkm","10,001-20,000 sqkm",paste("20,001-",max(wdpa_kfw$REP_AREA)," sqkm",sep="")))

# ----- load loss data and prepare for mapping -----
## loss statistics
gfw_lossstats <-
  read_csv(
    "../../datalake/mapme.protectedareas/output/polygon/global_forest_watch/zonal_statistics_allPAs_long_temporal.csv",
    col_types = c("cfid")
  )

## filter for area only
gfw_lossstats <-
  gfw_lossstats %>%
  filter(name == "area")

## load spatial data
wdpa_allPAs<-
  st_read("../../datalake/mapme.protectedareas/input/wdpa_kfw/wdpa_kfw_spatial_latinamerica_2021-04-22_allPAs_valid_simplified.gpkg")

wdpa_allPAs <- wdpa_allPAs %>%
  filter (ISO3 == "BOL")

## check if wdpa ids from loss statistics are contained in spatial data
table(unique(gfw_lossstats$WDPA_PID)%in%wdpa_allPAs$WDPA_PID) # all but one

## add column to spatial data that shows forest area in 2000 for each area (maximum is equal to area in 2000 since no forest gain is in the data)
max_forest <-
  gfw_lossstats %>%
  group_by(WDPA_PID) %>%
  dplyr::summarise(max_forest = max(value))

## merge to spatial and clean up
wdpa_allPAs <- 
  merge(wdpa_allPAs, max_forest,"WDPA_PID")

rm(max_forest)

## categorize maximum forest area by meaningfull and pretty breaks
wdpa_allPAs$max_forest_categorized <-
  cut(
    wdpa_allPAs$max_forest,
    breaks = c(-1, 100, 1000, 10000,100000, max(wdpa_allPAs$max_forest, na.rm = T)),
    labels = c("0-100", "101-1,000", "1,001-10,000", "10,001-100,000",">100,000")
  )

## calculate relative losses between 2000 and 2020 and add column to spatial data
relative_loss <-
  gfw_lossstats %>%
  group_by(WDPA_PID) %>%
  dplyr::summarise(relative_loss = 1-min(value)/max(value),absolute_loss = max(value)-min(value))

## merge to spatial and clean up
wdpa_allPAs <- 
  merge(wdpa_allPAs, relative_loss, "WDPA_PID")

rm(relative_loss)

## categorize relative loss by meaningfull and pretty breaks 
wdpa_allPAs$relative_loss_categorized <-
  cut(
    wdpa_allPAs$relative_loss,
    breaks = c(-1, 0.02, 0.05, 0.1, 0.2, 1),
    labels = c("0-2 %", "2-5 %", "5-10 %", "10-20 %",">20 %")
  )


## categorize absolute loss by meaningfull and pretty breaks 
wdpa_allPAs$absolute_loss_categorized <-
  cut(
    wdpa_allPAs$absolute_loss,
    breaks = c(-1, 10, 100, 1000, 10000,max(wdpa_allPAs$absolute_loss)),
    labels = c("0-10", "10-100", "100-1,000", "1,000-10,000",">10,000")
  )

## subset data and get centroids for map
wdpa_allPAs_lossdata<-wdpa_allPAs %>% 
  filter(DESIG_ENG != 'UNESCO-MAB Biosphere Reserve') %>% 
  filter(GEOMETRY_TYPE != 'POINT') %>% 
  filter(is.nan(relative_loss)!=TRUE) %>%
  st_centroid()

## subset all data for untreated
wdpa_nonSupportPA<-wdpa_allPAs %>% 
  filter(DESIG_ENG != 'UNESCO-MAB Biosphere Reserve') %>% 
  filter(GEOMETRY_TYPE != 'POINT') %>% 
  filter(is.na(bmz_n_1)==TRUE) 

## add fire data
bolivia_fire_data <- read.csv("data/bolivia_fires_2000-2021.csv")
names(bolivia_fire_data) [1] <- "WDPAID"
wdpa_allPAs_lossdata <- merge(wdpa_allPAs_lossdata, bolivia_fire_data, by = "WDPAID", all.x = TRUE)
wdpa_allPAs_lossdata$fires_categorized <-
  cut(
    wdpa_allPAs_lossdata$active_fire_counts,
    breaks = c(0, 500, 2500, 5000, 10000, 1e6),
    labels = c("0-500", "501-2,500", "2501-5,000", "5,001-10,000","> 10,000")
  )

bolivia_suggested_PAs <- wdpa_kfw %>% filter(WDPAID %in% bol_wdpa_ids)
```
## Introduction
The following analysis is intented to support a peer-discussion on the threat level of protected areas (short: PAs) in the Bolivian Amazon Basin. The goal of this exercise is to assist KfW and its partners in the project preparation phase. In addition, information about the current portfolio and its relevance to protect the most threatened areas can be derived.

The goal of this analysis is to look at a set of predefined PAs, rank it according to the their threat level and compare it to the past portfolio from KfW. The analysis is based on publicly available data from the World Database of Protected Areas (WDPA/IUCN) [^1] and other freely available geodata-sources. In a first analysis step we focus on habitat destruction in terms of primary forest cover loss and in the second step we look at fires in protected areas between 2000 and 2021. Both datasets can indicate human pressures on the ecosystem as well as natural/climatic stressors that could harm the long-term stability and provision of ecosystem services. 

## Forest cover loss (2000-2020)
To quantify forest cover loss we utilize data from the Global Forest Watch (Hansen et al, 2013)[^2]. Forest cover loss is defined in their methdology _"as a stand-replacement disturbance, or a change from a forest to non-forest state."_. Loss can either be the result of human activities or natural factors such as droughts, forest fires or hurricanes, amongst others. More information on the interpretation and usefullness of this data as well as suggested further steps to advance the threat assessment are given below in the discussion part. The data also include annual forest fire counts, which are taken from the NASA FIRMS data[^3].

__In the analysis  we will focus on two key outcome indicators:__ 

* **Total forest cover loss**: Measures the total sum of loss areas in hectare. This variable is able to identify PAs with the highest primary forest cover loss between 2001 and 2020. The identification of high loss areas can be usefull for targeting areas where we might achieve the largest impact in terms of _reducing emissions from deforstation and forest degradation_. 

* **Relative forest cover loss**: Measures the percentage of primary forest cover loss inside a PA compared to its total primary forest area in 2000. The identification of PAs with high relative losses can be relevant from a biodiversity perspective. PAs with high relative losses might be places where large parts of the functional forest habitat is lost. Targeting these areas might not only help to protect the _floral biodiversity but also the fauna and humans that inhabit these areas and profit from the local forest ecosystem services_. 

### Map
The following map depicts relative and absolute forest cover loss in the selected areas. The size of circles is dependent on the total loss (the bigger the total loss, the larger the circle). The color is dependend on the relative loss (red circles indicate areas with high loss). From a threat perspective big red circles could be especially relevant areas for conservation. This map is interactive meaning that you can zoom into the map to see specific countries in more detail and click on areas to get summary statistics. Furthermore, supported PAs from the current and past portfolio of KfW (blue) are displayed with their actual polygon boundaries as well as all other PAs in Bolivia (grey) which can be activated manually in the map. Furthermore the analyzed data from Global Forest watch can be seen when activated in the layer control panel as well as the distribution of primary forests in 2001.

```{r projectmap_def, echo=FALSE, warning=FALSE,  fig.width=10}
## ----- create color palettes for the map -----
pal_relative_loss <- colorFactor(
  palette = rev(brewer.pal(5, "RdYlGn")),
  domain = wdpa_allPAs$relative_loss_categorized
)

pal_absoute_loss <- colorFactor(
  palette = rev(brewer.pal(5, "RdYlGn")),
  domain = wdpa_allPAs$absolute_loss_categorized
)

# create colorramp function for country
colourCount = length(unique(wdpa_kfw$ISO3))
getPalette = colorRampPalette(brewer.pal(9, "Set1"))

## add label string for map display 
wdpa_allPAs_lossdata$label <- 
  with(wdpa_allPAs_lossdata, paste(
  "<p> <b>", NAME, "</b> </br>",
  "Primary Forests (2000)",round(max_forest,digits=0)," ha", "</br>",
  "Forest Fires (2000-2021)", active_fire_counts, "</br>",
  "Absolute Loss (2000-2020):",round(absolute_loss,digits=0)," ha", "</br>",
  "Relative Loss (2000-2020):", round(relative_loss*100,digits=2)," %:",
  "</p>"))

# subset data for treated only 
wdpa_kfw_treated<-wdpa_kfw %>% 
  filter(is.na(bmz_n_1)==F)

# subset lossdata for selected only
wdpa_allPAs_lossdata_subset<-wdpa_allPAs_lossdata %>% 
  filter(WDPAID %in% bol_wdpa_ids)

# get centroid coordinates
wdpa_allPAs_lossdata_subset<-
  cbind(wdpa_allPAs_lossdata_subset,
        st_coordinates(st_centroid(wdpa_allPAs_lossdata_subset)))

# ----- create map -----
my_map <-
  leaflet() %>%
  # add external map providers
  addTiles(group = "OpenStreetMap") %>%
  addProviderTiles(providers$CartoDB.Positron, group="CartoDB") %>%
  addProviderTiles(providers$Esri.WorldImagery, group="Satellite") %>%
  addProviderTiles(providers$Esri.WorldShadedRelief, group="Topography") %>%
  addProviderTiles(providers$NASAGIBS.ViirsEarthAtNight2012, group="Nightlights") %>%
  addTiles(
    "https://tiles.globalforestwatch.org/umd_tree_cover_loss/latest/dynamic/{z}/{x}/{y}.png",
    group = "Forest Cover Loss (2001-2020)",
    #options=tileOptions(opacity = 0.7),
    attribution = "Hansen, M. C., P. V. Potapov, R. Moore, M. Hancher, S. A. Turubanova, A. Tyukavina, D. Thau, S. V. Stehman, S. J. Goetz, T. R. Loveland, A. Kommareddy, A. Egorov, L. Chini, C. O. Justice, and J. R. G. Townshend. 2013. “High-Resolution Global Maps of 21st-Century Forest Cover Change.” Science 342 (15 November): 850–53. Data available on-line from: http://earthenginepartners.appspot.com/science-2013-global-forest."
  ) %>%
  addTiles(
    "https://tiles.globalforestwatch.org/umd_regional_primary_forest_2001/latest/dynamic/{z}/{x}/{y}.png",
    group = "Regional Primary Forests (2001)",
    attribution = "Hansen, M. C., P. V. Potapov, R. Moore, M. Hancher, S. A. Turubanova, A. Tyukavina, D. Thau, S. V. Stehman, S. J. Goetz, T. R. Loveland, A. Kommareddy, A. Egorov, L. Chini, C. O. Justice, and J. R. G. Townshend. 2013. “High-Resolution Global Maps of 21st-Century Forest Cover Change.” Science 342 (15 November): 850–53. Data available on-line from: http://earthenginepartners.appspot.com/science-2013-global-forest."
  ) %>%
  # add own polygon data
  addPolygons(data = wdpa_kfw_treated,opacity = 0.9,color = "#2b8cbe", group = "PAs KfW-supported",label = ~htmlEscape(NAME),weight = 1)%>%
  addPolygons(data = wdpa_nonSupportPA,opacity = 0.9,color = "darkgrey", group = "PAs (Others)",label = ~htmlEscape(NAME),weight = 2)%>%
  addPolygons(data = bolivia_suggested_PAs,opacity = 1, fillOpacity = 0.1, color = "#7d0f02", group = "PAs Suggested",label = ~htmlEscape(NAME),weight = 2)%>%
  # add own forest loss data as circles
  addCircleMarkers(data=wdpa_allPAs_lossdata_subset,
                   color = ~pal_relative_loss(relative_loss_categorized),
                   group = "Sum of Forest Cover Loss (2001-2020)",
                   radius = ~ifelse(absolute_loss_categorized == ">10,000", 12, 
                                    ifelse(absolute_loss_categorized == "1,000-10,000", 6,
                                           ifelse(absolute_loss_categorized == "100-1,000", 4,
                                                  ifelse(absolute_loss_categorized == "10-100", 2,1
                                                  )      
                                           )
                                    )),
                   popup = ~label,
                   stroke = FALSE, 
                   fillOpacity = 0.5
                   ) %>%
    # labels
  addLabelOnlyMarkers(data = wdpa_allPAs_lossdata_subset,
                    lng = ~X, lat = ~Y, label = ~NAME,group = "Labels (PA Names)",
                    labelOptions = labelOptions(noHide = TRUE, direction = 'top', textOnly = TRUE)) %>% 
  # add fullscreen control
  addFullscreenControl() %>%
  # add legend(s)
  addLegend("bottomright",
            data = wdpa_allPAs,
            pal = pal_relative_loss,
            values = ~relative_loss_categorized,
            title = "Relative Forest Cover Loss (2001-2020)",
            opacity = 1,
            group = "Sum of Forest Cover Loss (2001-2020)") %>%
  
  # add layers control to define how and where data is displayed.
  addLayersControl(
    baseGroups = c("CartoDB","OpenStreetMap","Satellite","Topography","Nightlights"), #"Toner",,"Regional Primary Forests (2001)"
    overlayGroups = c("PAs Suggested", "PAs KfW-supported","PAs (Others)", "Labels (PA Names)", "Sum of Forest Cover Loss (2001-2020)","Forest Cover Loss (2001-2020)", "Regional Primary Forests (2001)"),
    options = layersControlOptions(collapsed = FALSE)) %>%
  # ommit certain layers
  hideGroup(group = c("PAs (Others)","Forest Cover Loss (2001-2020)", "Regional Primary Forests (2001)","Labels (PA Names)"))

my_map
```

### Absolute Trend
Lollipop Plots can be used to compare PAs in terms of their treat level. The following figure shows absolut forest cover loss in suggested PAs. It also shows whether any of these PAs has been or is currently supported by a project from KfW. You can isolate single areas with a double click on their name in the legend. 

```{r lollipop_plot-suggested-regions-fcl, message = FALSE, warning = FALSE,fig.width=10, echo=FALSE, eval = TRUE}
# ----- create horizontal lollipop plots for top loss areas -----
lollipop_data <- wdpa_allPAs_lossdata %>%
  filter(WDPAID %in% bol_wdpa_ids)

lollipop_data$display_name <- ifelse (is.na(lollipop_data$bmz_n_1),
        paste0(lollipop_data$NAME),
        paste0(lollipop_data$NAME, " (", lollipop_data$bmz_n_1, ")")
        )

lossdata_lollipop_plot <-
  lollipop_data %>%
  # arrange(desc(absolute_loss)) %>%
  ggplot() +
  geom_point(aes(
    x = reorder(display_name,-absolute_loss),
    y = absolute_loss,
    color = factor(bmz_n_1),
    text = paste(
      "Area Name: ", display_name,
      "\nForest cover loss: ", round(absolute_loss,digits = 0), " ha",
      "\nBMZ-Number: ", bmz_n_1)
  )) +
  geom_segment(aes(
    x = reorder(display_name,-absolute_loss),
    xend = reorder(display_name,-absolute_loss),
    y = 0,
    yend = absolute_loss,
    color = factor(bmz_n_1),
    text = paste(
      "Area Name: ", display_name,
      "\nForest cover loss: ", round(absolute_loss,digits = 0), " ha",
      "\nBMZ-Number: ", bmz_n_1)
  )) +
  coord_flip() +
  labs(x="", y="Absolute Forest Cover Loss in ha (2000-2020)", color = "Project Number")+
  theme_classic()

ggplotly(lossdata_lollipop_plot,tooltip = "text")
```

### Yearly Forest Cover Loss
Stacked Barplots are used to assess the time-trend of one or several PAs. They show the variation in the total number of fires which can be amongst others attributed to El Niño and La Niña which are opposite extremes of ENSO, referring to cyclical environmental conditions that occur across the Equatorial Pacific Ocean. For more information see [here](https://www.noaa.gov/education/resource-collections/weather-atmosphere/el-nino). The different parts of each bar represent the different PAs. You can easily see which PAs take a larger share of the yearly total fire counts. Those are the ones where most fires occur. You can get more detailed information by hovering with the mouse over the bars. 

```{r line_plots_forestcoverloss, message = FALSE, warning = FALSE,fig.width=10, echo=FALSE, eval=TRUE}
bol_annual_fires <- read.csv("data/bolivia_annual_fires_2000_2021.csv")
fire_lines_aux_data <- st_drop_geometry(wdpa_kfw) %>%
  select(c(WDPAID, bmz_n_1, NAME))
names(fire_lines_aux_data) [1] <- "wdpa_id"
bol_annual_fires <- merge (bol_annual_fires, fire_lines_aux_data, by = "wdpa_id", all.x = TRUE)

palette_length <- length(unique(bol_annual_fires$bmz_n_1))
palette_lineplot<-
  c(brewer.pal(palette_length-1,"Set1"),"grey80")

# create lineplot
fires_absolute_lineplot <-
  bol_annual_fires %>%
  ggplot(.,aes(x=year,
                y=active_fire_counts,
                fill=factor(NAME))) +
  geom_bar(position="stack", 
           stat="identity")+
  labs(x="",title="Annual Forest Fires in Suggested PAs")+
  scale_y_continuous(labels=function(x) format(x, big.mark = ",", scientific = FALSE),name="Forest Loss in ha")+
  theme_classic()

ggplotly(fires_absolute_lineplot)
```

## Fire counts (2000 to 2021)
### Map
The following map depicts a summary of fire counts in the suggested Protected Areas according to NASA Firms. In general larger protected areas expirience more fires due to their spatial extension. Nevertheless there are also exeptions. It appears that large protected areas in the Andean mountain range expirience less forest fires (you can activate a Topography layer in the map). In contrast, PAs in the east which are located closer to the Brazilian boarder (Rondonia) expirience more fires. There is also a notable correlation between forest cover loss and forest fires (You can activate the forest cover loss layer in the map). This seems plausible since fires are often used as a strategy for forest cleansing or large natural wildfires might cause permanent damage to the forest cover. However, the relationship is not perfect. 
The map and the subsequent lollipop plots also exhibit which of the analyzed areas have been already part of past or ongoing KfW projects. The data suggests that only one area "Noel Kempff Mercado" expirienced a larger amount of fires (6,724) between 2000 and 2021. You can get individual area statistics by clicking on the colored circles in the map. 

```{r projectmap_fire, echo=FALSE, warning=FALSE,  fig.width=10}
## ----- create color palettes for the map -----
pal_fires <- colorFactor(
  palette = rev(brewer.pal(5, "RdYlGn")),
  domain = wdpa_allPAs_lossdata$fires_categorized
)

## add label string for map display 
wdpa_allPAs_lossdata$label <- 
  with(wdpa_allPAs_lossdata, paste(
  "<p> <b>", NAME, "</b> </br>",
  "Primary Forests (2000)",round(max_forest,digits=0)," ha", "</br>",
  "Forest Fires (2000-2021)", active_fire_counts, "</br>",
  "Absolute Loss (2000-2020):",round(absolute_loss,digits=0)," ha", "</br>",
  "Relative Loss (2000-2020):", round(relative_loss*100,digits=2)," %:",
  "</p>"))

# subset data for treated only 
wdpa_kfw_treated<-wdpa_kfw %>% 
  filter(is.na(bmz_n_1)==F)

# subset lossdata for selected only
wdpa_allPAs_lossdata_subset<-wdpa_allPAs_lossdata %>% 
  filter(WDPAID %in% bol_wdpa_ids)

# get centroid coordinates
wdpa_allPAs_lossdata_subset<-
  cbind(wdpa_allPAs_lossdata_subset,
        st_coordinates(st_centroid(wdpa_allPAs_lossdata_subset)))

# ----- create map -----
my_map <-
  leaflet() %>%
  # add external map providers
  addTiles(group = "OpenStreetMap") %>%
  addProviderTiles(providers$CartoDB.Positron, group="CartoDB") %>%
  addProviderTiles(providers$Esri.WorldImagery, group="Satellite") %>%
  addProviderTiles(providers$Esri.WorldShadedRelief, group="Topography") %>%
  addProviderTiles(providers$NASAGIBS.ViirsEarthAtNight2012, group="Nightlights") %>%
  addTiles(
    "https://tiles.globalforestwatch.org/umd_tree_cover_loss/latest/dynamic/{z}/{x}/{y}.png",
    group = "Forest Cover Loss (2001-2020)",
    #options=tileOptions(opacity = 0.7),
    attribution = "Hansen, M. C., P. V. Potapov, R. Moore, M. Hancher, S. A. Turubanova, A. Tyukavina, D. Thau, S. V. Stehman, S. J. Goetz, T. R. Loveland, A. Kommareddy, A. Egorov, L. Chini, C. O. Justice, and J. R. G. Townshend. 2013. “High-Resolution Global Maps of 21st-Century Forest Cover Change.” Science 342 (15 November): 850–53. Data available on-line from: http://earthenginepartners.appspot.com/science-2013-global-forest."
  ) %>%
  addTiles(
    "https://tiles.globalforestwatch.org/umd_regional_primary_forest_2001/latest/dynamic/{z}/{x}/{y}.png",
    group = "Regional Primary Forests (2001)",
    attribution = "Hansen, M. C., P. V. Potapov, R. Moore, M. Hancher, S. A. Turubanova, A. Tyukavina, D. Thau, S. V. Stehman, S. J. Goetz, T. R. Loveland, A. Kommareddy, A. Egorov, L. Chini, C. O. Justice, and J. R. G. Townshend. 2013. “High-Resolution Global Maps of 21st-Century Forest Cover Change.” Science 342 (15 November): 850–53. Data available on-line from: http://earthenginepartners.appspot.com/science-2013-global-forest."
  ) %>%
  # add own polygon data
  addPolygons(data = wdpa_kfw_treated,opacity = 0.9,color = "#2b8cbe", group = "PAs KfW-supported",label = ~htmlEscape(NAME),weight = 1)%>%
  addPolygons(data = wdpa_nonSupportPA,opacity = 0.9,color = "darkgrey", group = "PAs (Others)",label = ~htmlEscape(NAME),weight = 2)%>%
  addPolygons(data = bolivia_suggested_PAs,opacity = 1, fillOpacity = 0.1, color = "#7d0f02", group = "PAs Suggested",label = ~htmlEscape(NAME),weight = 2)%>%
  # add own active fire counts data as circles
  addCircleMarkers(data=wdpa_allPAs_lossdata_subset,
                   color = ~pal_fires(fires_categorized),
                   group = "Active Fires Count (2000-2021)",
                   popup = ~label,
                   stroke = FALSE, 
                   fillOpacity = 0.8
                   ) %>%
  # labels
  addLabelOnlyMarkers(data = wdpa_allPAs_lossdata_subset,
                    lng = ~X, lat = ~Y, label = ~NAME,group = "Labels (PA Names)",
                    labelOptions = labelOptions(noHide = TRUE, direction = 'top', textOnly = TRUE)) %>% 
  # controls
  addFullscreenControl() %>%
  # add legend(s)
  addLegend("bottomright",
            data = wdpa_allPAs_lossdata,
            pal = pal_fires,
            values = ~fires_categorized,
            title = "Count of Active Fires in PAs (2000-2021)",
            opacity = 1,
            group = "Active Fires (2000-2021)") %>%
  
  # add layers control to define how and where data is displayed.
  addLayersControl(
    baseGroups = c("CartoDB","OpenStreetMap","Satellite","Topography","Nightlights"), #"Toner",,"Regional Primary Forests (2001)"
    overlayGroups = c("PAs Suggested", "PAs KfW-supported","PAs (Others)","Labels (PA Names)", "Active Fires Count (2000-2021)","Forest Cover Loss (2001-2020)", "Regional Primary Forests (2001)"),
    options = layersControlOptions(collapsed = FALSE)) %>%
  # ommit certain layers
  hideGroup(group = c("PAs (Others)","PAs KfW-supported","Forest Cover Loss (2001-2020)", "Regional Primary Forests (2001)","Labels (PA Names)"))

my_map
```

### Absolute Trend
Lollipop Plots can be used to compare PAs in terms of their treat level. The following figure shows forest fires in suggested PAs. It also shows whether any of these PAs has been or is currently supported by a project from KfW. <

```{r lollipop_plot-suggested-regions, message = FALSE, warning = FALSE,fig.width=10, echo=FALSE, eval = TRUE}
# ----- create horizontal lollipop plots for top loss areas -----
lollipop_data <- wdpa_allPAs_lossdata %>%
  filter(WDPAID %in% bol_wdpa_ids)

lollipop_data$display_name <- ifelse (is.na(lollipop_data$bmz_n_1),
        paste0(lollipop_data$NAME),
        paste0(lollipop_data$NAME, " (", lollipop_data$bmz_n_1, ")")
        )

lossdata_lollipop_plot <-
  lollipop_data %>%
  # arrange(desc(absolute_loss)) %>%
  ggplot() +
  geom_point(aes(
    x = reorder(display_name,-active_fire_counts),
    y = active_fire_counts,
    color = factor(bmz_n_1),
    text = paste(
      "Area Name: ", display_name,
      "\nCount of Active Fires: ", round(active_fire_counts,digits = 0), " fires",
      "\nBMZ-Number: ", bmz_n_1)
  )) +
  geom_segment(aes(
    x = reorder(display_name,-active_fire_counts),
    xend = reorder(display_name,-active_fire_counts),
    y = 0,
    yend = active_fire_counts,
    color = factor(bmz_n_1),
    text = paste(
      "Area Name: ", display_name,
      "\nCount of Active Fires: ", round(active_fire_counts,digits = 0), " fires",
      "\nBMZ-Number: ", bmz_n_1)
  )) +
  coord_flip() +
  labs(x="", y="Count of Fires (2000-2021)", color = "Project Number")+
  theme_classic()

ggplotly(lossdata_lollipop_plot,tooltip = "text")
```
  
### Yearly Firecounts
Stacked Barplots are used to assess the time-trend of one or several PAs. They show the variation in the total number of fires which can be amongst others attributed to El Niño and La Niña which are opposite extremes of ENSO, referring to cyclical environmental conditions that occur across the Equatorial Pacific Ocean. For more information see [here](https://www.noaa.gov/education/resource-collections/weather-atmosphere/el-nino). The different parts of each bar represent the different PAs. You can easily see which PAs take a larger share of the yearly total fire counts. Those are the ones where most fires occur. You can get more detailed information by hovering with the mouse over the bars. 


```{r line_plots-fire, message = FALSE, warning = FALSE,fig.width=10, echo=FALSE, eval=TRUE}
bol_annual_fires <- read.csv("data/bolivia_annual_fires_2000_2021.csv")
fire_lines_aux_data <- st_drop_geometry(wdpa_kfw) %>%
  select(c(WDPAID, bmz_n_1, NAME))
names(fire_lines_aux_data) [1] <- "wdpa_id"
bol_annual_fires <- merge (bol_annual_fires, fire_lines_aux_data, by = "wdpa_id", all.x = TRUE)

palette_length <- length(unique(bol_annual_fires$bmz_n_1))
palette_lineplot<-
  c(brewer.pal(palette_length-1,"Set1"),"grey80")

# create lineplot
fires_absolute_lineplot <-
  bol_annual_fires %>%
  ggplot(.,aes(x=year,
                y=active_fire_counts,
                # group=factor(NAME)
                fill=factor(NAME))) +
  geom_bar(position="stack", 
           stat="identity")+
  labs(x="",title="Annual Forest Fires in Suggested PAs")+
  scale_y_continuous(labels=function(x) format(x, big.mark = ",", scientific = FALSE),name="Forest Loss in ha")+
  #scale_color_manual(name = "Project Number", values = palette_lineplot)+
  theme_classic()

ggplotly(fires_absolute_lineplot)
```

## Interpretation & Discussion 

### Forest Loss Data

As outline above the GFW methodology defines forest cover loss as a _"as a stand-replacement disturbance, or a change from a forest to non-forest state."_. Loss can either be the result of human activities or natural factors such as droughts, forest fires or hurricanes, amongst others. Thus, the data currently does not allow the differentiate permanent loss and conversion from temporary loss due to natural factors. **A more in-depth analysis of pre-selected areas is therefore recommended when using the current data.**

GFW data can be used to get an assessment of old-growth and primary forests as well as associated carbon emissions. It can indicate which areas had been highly threatened either due to natural causes (hurricanes/fires) or human causes (deforestation/logging/degradation and conversion to agriculture/silviculture). Especially useful in this context is to have a look at the whole trend from 2001 to 2020. Huge spikes in the individual trend data might indicate natural causes such as fires and hurricanes. A quick search with google always helps to confirm this hypothesis or you might look at Hurricane data e.g. on the [NOAA website](https://coast.noaa.gov/hurricanes/#map=4/32/-80) which provides historical data for hurricane tracks. More continuous growth in forest loss is probably due to the conversion of natural forests for agricultural purposes. Again you might look at the map given above and activate the satellite layer to see what might be happening below the detected loss. GFW does not (currently) allow to detect regrowth and regeneration but it is planned to provide that feature soon.  

**Possible Improvements**

* Analyze subsequent land-cover in loss-areas. This would be possible e.g. by using Landcover data from the ESA Copernicus mission. It would allow to seperate permanent conversion of forest area e.g. for agricultural purposes from natural losses e.g. due wildfires.
* Aggregate existing loss data on a finer spatial grid or create heatmaps. This would allow to better visualize highly threatened  spots on the map. This could be especially relevant to quickly identify high pressure zones in large protected areas. 
* Analyse bufferzones: This would allow to detect areas with high forest loss dynamics in its sourroundings. 


[comment]: <> (Literature)

[^1]: UNEP-WCMC and IUCN (2022), Protected Planet: The World Database on Protected Areas (WDPA) and World Database on Other Effective Area-based Conservation Measures (WD-OECM) [Online], February 2022, Cambridge, UK: UNEP-WCMC and IUCN. Available at: www.protectedplanet.net.

[^2]: "Hansen, M. C., P. V. Potapov, R. Moore, M. Hancher, S. A. Turubanova, A. Tyukavina, D. Thau, S. V. Stehman, S. J. Goetz, T. R. Loveland, A. Kommareddy, A. Egorov, L. Chini, C. O. Justice, and J. R. G. Townshend. 2013. “High-Resolution Global Maps of 21st-Century Forest Cover Change.” Science 342 (15 November): 850–53. Data available on-line from: http://earthenginepartners.appspot.com/science-2013-global-forest."

[^3]: "Justice, C.O., Giglio, L., Korontzi, S., Owens, J., Morisette, J., Roy, D., Descloitres, J., Alleaume, S., Petitcolin, F., & Kaufman, Y.J. (2002). The MODIS fire products. Remote Sensing of Environment, 83:244-262. doi:10.1016/S0034-4257(02)00076-7"