---
title: "Project Preparation: Threat Assessment for Protected Areas"
author: "Johannes Schielein (Author)"
date: "2022-02-10"
output: workflowr::wflow_html
editor_options:
  chunk_output_type: console
---

```{r setup, message = FALSE, warning = FALSE, include = FALSE}
# load libraries and set session options. 
library("tidyverse")
library("sf")
library("leaflet")
library("leaflet.extras")
library("leaflet.extras2")
library("ggsci")
library("scales")
library("htmltools")
library("RColorBrewer")
library("plotly")
options(scipen=10000)
```

```{r database_creation, message = FALSE, warning = FALSE, include = FALSE}
# ----- load and transform protected areas data -----
##  Protected areas
wdpa_kfw<-
  read_sf("~/shared/datalake/mapme.protectedareas/input/wdpa_kfw/wdpa_kfw_spatial_latinamerica_2021-02-01_supportedPAs_unique_simplified.gpkg")

## create column for area coloring based on categories
wdpa_kfw$REP_AREA_cat<-
  cut(wdpa_kfw$REP_AREA,
      c(0,1000,5000,10000,20000,max(wdpa_kfw$REP_AREA)),
      c("< 1,000 sqkm","1,001-5,000 sqkm","5,001-10,000 sqkm","10,001-20,000 sqkm",paste("20,001-",max(wdpa_kfw$REP_AREA)," sqkm",sep="")))

# ----- load loss data and prepare for mapping -----
## loss statistics
gfw_lossstats <-
  read_csv(
    "../../datalake/mapme.protectedareas/output/polygon/global_forest_watch/zonal_statistics_allPAs_long_temporal.csv",
    col_types = c("cfid")
  )

## filter for area only
gfw_lossstats <-
  gfw_lossstats %>%
  filter(name == "area")

## load spatial data
wdpa_allPAs<-
  st_read("../../datalake/mapme.protectedareas/input/wdpa_kfw/wdpa_kfw_spatial_latinamerica_2021-04-22_allPAs_valid_simplified.gpkg")

## check if wdpa ids from loss statistics are contained in spatial data
table(unique(gfw_lossstats$WDPA_PID)%in%wdpa_allPAs$WDPA_PID) # all but one

## add column to spatial data that shows forest area in 2000 for each area (maximum is equal to area in 2000 since no forest gain is in the data)
max_forest <-
  gfw_lossstats %>%
  group_by(WDPA_PID) %>%
  dplyr::summarise(max_forest = max(value))

## merge to spatial and clean up
wdpa_allPAs <- 
  merge(wdpa_allPAs, max_forest,"WDPA_PID")

rm(max_forest)

## categorize maximum forest area by meaningfull and pretty breaks
wdpa_allPAs$max_forest_categorized <-
  cut(
    wdpa_allPAs$max_forest,
    breaks = c(-1, 100, 1000, 10000,100000, max(wdpa_allPAs$max_forest, na.rm = T)),
    labels = c("0-100", "101-1,000", "1,001-10,000", "10,001-100,000",">100,000")
  )

## calculate relative losses between 2000 and 2020 and add column to spatial data
relative_loss <-
  gfw_lossstats %>%
  group_by(WDPA_PID) %>%
  dplyr::summarise(relative_loss = 1-min(value)/max(value),absolute_loss = max(value)-min(value))

## merge to spatial and clean up
wdpa_allPAs <- 
  merge(wdpa_allPAs, relative_loss, "WDPA_PID")

rm(relative_loss)

## categorize relative loss by meaningfull and pretty breaks 
wdpa_allPAs$relative_loss_categorized <-
  cut(
    wdpa_allPAs$relative_loss,
    breaks = c(-1, 0.02, 0.05, 0.1, 0.2, 1),
    labels = c("0-2 %", "2-5 %", "5-10 %", "10-20 %",">20 %")
  )


## categorize absolute loss by meaningfull and pretty breaks 
wdpa_allPAs$absolute_loss_categorized <-
  cut(
    wdpa_allPAs$absolute_loss,
    breaks = c(-1, 10, 100, 1000, 10000,max(wdpa_allPAs$absolute_loss)),
    labels = c("0-10", "10-100", "100-1,000", "1,000-10,000",">10,000")
  )

## subset data and get centroids for map
wdpa_allPAs_lossdata<-wdpa_allPAs %>% 
  filter(DESIG_ENG != 'UNESCO-MAB Biosphere Reserve') %>% 
  filter(GEOMETRY_TYPE != 'POINT') %>% 
  filter(is.nan(relative_loss)!=TRUE) %>%
  st_centroid()

## subset all data for untreated
wdpa_nonSupportPA<-wdpa_allPAs %>% 
  filter(DESIG_ENG != 'UNESCO-MAB Biosphere Reserve') %>% 
  filter(GEOMETRY_TYPE != 'POINT') %>% 
  filter(is.na(bmz_n_1)==TRUE) 

```
## Introduction
The following graphics are intented to support a peer-discussion on how publicly available geodata can be used to characterize the threat level of protected areas (short PAs). The goal of this exercise is to define data products that could support the identification of new intervention sites and thus help KfW and its partners in the project preparation phase. In addition, information about the current portfolio and its relevance to protect the most threatened areas can be derived.

The general idea of this analysis is to match the complete network of existing PAs in countries where KfW is active to the the database of supported PAs by KfW. The same methodology can be used later in the project preparation phase to analyse the existing portfolio together with newly projected intervention sites. The analysis is based on publicly available data from the World Database of Protected Areas (WDPA/IUCN).[^1] Our current portfolio database contains `r nrow(wdpa_allPAs[!is.na(wdpa_allPAs$bmz_n_1),])` areas from a total of `r nrow(wdpa_allPAs)` in the partnering countries. 

In a first analysis step we focus on habitat destruction in terms of primary forest cover loss. To quantify forest cover loss we utilize data from the Global Forest Watch (Hansen et al, 2013)[^2]. Forest cover loss is defined in their work _"as a stand-replacement disturbance, or a change from a forest to non-forest state."_. Loss can either be the result of human activities or natural factors such as droughts, forest fires or hurricanes, amongst others. More information on the interpretation and usefullness of this data as well as suggested further steps to advance the threat assessment are given below in the discussion part. 

__In the analysis below we will focus on two key outcome indicators:__ 

* **Total forest cover loss**: Measures the total sum of loss areas in hectare. This variable is able to identify PAs with the highest primary forest cover loss between 2001 and 2020. The identification of high loss areas can be usefull for targeting areas where we might achieve the largest impact in terms of _reducing emissions from deforstation and forest degradation_. 

* **Relative forest cover loss**: Measures the percentage of primary forest cover loss inside a PA compared to its total primary forest area in 2000. The identification of PAs with high relative losses can be relevant from a biodiversity perspective. PAs with high relative losses might be places where large parts of the functional forest habitat is lost. Targeting these areas might not only help to protect the _floral biodiversity but also the fauna and humans that inhabit these areas and profit from the local forest ecosystem services_. 

## Map of Threatened Areas
The following map depicts relative and absolute forest cover loss for PAs in KfW supported areas. The size of circles is dependent on the total loss (the bigger the total loss, the larger the circle). The color is dependend on the relative loss (red circls indicate areas with high loss). From a threat perspective big red circles could be especially relevant areas for conservation. You can also zoom into the map to see the PA boundaries and which areas are in the portfolio of KfW (blue) and which are not (grey). 

```{r projectmap, echo=FALSE, warning=FALSE,  fig.width=10}
## ----- create color palettes for the map -----
pal_relative_loss <- colorFactor(
  palette = rev(brewer.pal(5, "RdYlGn")),
  domain = wdpa_allPAs$relative_loss_categorized
)

pal_absoute_loss <- colorFactor(
  palette = rev(brewer.pal(5, "RdYlGn")),
  domain = wdpa_allPAs$absolute_loss_categorized
)

# create colorramp function for country
colourCount = length(unique(wdpa_kfw$ISO3))
getPalette = colorRampPalette(brewer.pal(9, "Set1"))

pal_country <- colorFactor(
  palette = getPalette(colourCount),
  domain = wdpa_kfw$ISO3
)

# create colorramp2
# pal_treatment <- colorFactor(
#   palette = c("darkblue","orange"),
#   domain = matched_data_merged$treat_ever
# )

## add label string for map display 
wdpa_allPAs_lossdata$label <- 
  with(wdpa_allPAs_lossdata, paste(
  "<p> <b>", NAME, "</b> </br>",
  "Primary Forests (2000)",round(max_forest,digits=0)," ha", "</br>",
  "Absolute Loss (2000-2020):",round(absolute_loss,digits=0)," ha", "</br>",
  "Relative Loss (2000-2020):", round(relative_loss*100,digits=2)," %:",
  "</p>"))

# ----- create map -----
my_map <-
  leaflet() %>%
  # add external map providers
  addTiles(group = "OpenStreetMap") %>%
  addProviderTiles(providers$CartoDB.Positron, group="CartoDB") %>%
  addProviderTiles(providers$Esri.WorldImagery, group="Satellite") %>%
  addProviderTiles(providers$Esri.WorldShadedRelief, group="Topography") %>%
  addProviderTiles(providers$NASAGIBS.ViirsEarthAtNight2012, group="Nightlights") %>%
  addTiles(
    "https://tiles.globalforestwatch.org/umd_tree_cover_loss/latest/dynamic/{z}/{x}/{y}.png",
    group = "Forest Cover Loss (2001-2020)",
    #options=tileOptions(opacity = 0.7),
    attribution = "Hansen, M. C., P. V. Potapov, R. Moore, M. Hancher, S. A. Turubanova, A. Tyukavina, D. Thau, S. V. Stehman, S. J. Goetz, T. R. Loveland, A. Kommareddy, A. Egorov, L. Chini, C. O. Justice, and J. R. G. Townshend. 2013. “High-Resolution Global Maps of 21st-Century Forest Cover Change.” Science 342 (15 November): 850–53. Data available on-line from: http://earthenginepartners.appspot.com/science-2013-global-forest."
  ) %>%
  addTiles(
    "https://tiles.globalforestwatch.org/umd_regional_primary_forest_2001/latest/dynamic/{z}/{x}/{y}.png",
    group = "Regional Primary Forests (2001)",
    attribution = "Hansen, M. C., P. V. Potapov, R. Moore, M. Hancher, S. A. Turubanova, A. Tyukavina, D. Thau, S. V. Stehman, S. J. Goetz, T. R. Loveland, A. Kommareddy, A. Egorov, L. Chini, C. O. Justice, and J. R. G. Townshend. 2013. “High-Resolution Global Maps of 21st-Century Forest Cover Change.” Science 342 (15 November): 850–53. Data available on-line from: http://earthenginepartners.appspot.com/science-2013-global-forest."
  ) %>%
  # add own polygon data
  addPolygons(data = wdpa_kfw,opacity = 0.9,color = "#2b8cbe", group = "PAs KfW-supported",label = ~htmlEscape(NAME),weight = 1)%>%
  addPolygons(data = wdpa_nonSupportPA,opacity = 0.7,color = "darkgrey", group = "PAs (Others)",label = ~htmlEscape(NAME),weight = 1)%>%
  # add own forest loss data as circles
  addCircleMarkers(data=wdpa_allPAs_lossdata,
                   color = ~pal_relative_loss(relative_loss_categorized),
                   group = "Relative Forest Loss (2001-2020)",
                   radius = ~ifelse(absolute_loss_categorized == ">10,000", 12, 
                                    ifelse(absolute_loss_categorized == "1,000-10,000", 6,
                                           ifelse(absolute_loss_categorized == "100-1,000", 4,
                                                  ifelse(absolute_loss_categorized == "10-100", 2,1
                                                  )      
                                           )
                                    )),
                   popup = ~label,
                   stroke = FALSE, 
                   fillOpacity = 0.5
                   ) %>%
  # add fullscreen control
  addFullscreenControl() %>%
  # add legend(s)
  addLegend("bottomright",
            data = wdpa_allPAs,
            pal = pal_relative_loss,
            values = ~relative_loss_categorized,
            title = "Relative Forest Loss (2001-2020)",
            opacity = 1,
            group = "PA Area Size") %>%
  
  # add layers control to define how and where data is displayed.
  addLayersControl(
    baseGroups = c("CartoDB","OpenStreetMap","Satellite","Topography","Nightlights"), #"Toner",,"Regional Primary Forests (2001)"
    overlayGroups = c("PAs KfW-supported","PAs (Others)","Relative Forest Loss (2001-2020)","Forest Cover Loss (2001-2020)", "Regional Primary Forests (2001)"),
    options = layersControlOptions(collapsed = FALSE)) %>%
  # ommit certain layers
  hideGroup(group = c("PA Area Size","Forest Cover Loss (2001-2020)", "Regional Primary Forests (2001)"))

my_map


```

## Lollipop plot 
Lollipop Plots can be used to compare supported PAs to non-supported PAs in terms of their treat level. The following figure shows the top 30 PAs that expirienced most forest cover loss in countries where KfW is active. It also shows whether any of these PAs has been or is currently supported by a project from KfW. The database also allows to filter for a specific country or to adapt the observed period (currently 2001-2020) or variable (relative loss instead of absolute). Interactive filtering could be enabled once an interactive solution (Dashboard) is in place. 

```{r lollipop_plot, message = FALSE, warning = FALSE,fig.width=10, echo=FALSE}
# ----- create horizontal lollipop plots for top loss areas -----
lossdata_lollipop_plot <-
  wdpa_allPAs_lossdata %>%
  top_n(30, wdpa_allPAs_lossdata$absolute_loss) %>%
  # arrange(desc(absolute_loss)) %>%
  ggplot() +
  geom_point(aes(
    x = reorder(NAME,-absolute_loss),
    y = absolute_loss,
    color = factor(bmz_n_1),
    text = paste(
      "Area Name: ", NAME,
      "\nForest cover loss: ", round(absolute_loss,digits = 0), " ha",
      "\nBMZ-Number: ", bmz_n_1)
  )) +
  geom_segment(aes(
    x = reorder(NAME,-absolute_loss),
    xend = reorder(NAME,-absolute_loss),
    y = 0,
    yend = absolute_loss,
    color = factor(bmz_n_1),
    text = paste(
      "Area Name: ", NAME,
      "\nForest cover loss: ", round(absolute_loss,digits = 0), " ha",
      "\nBMZ-Number: ", bmz_n_1)
  )) +
  coord_flip() +
  labs(x="", y="Absolute Forest Cover Loss in ha (2000-2020)", color = "Project Number")+
  theme_classic()

ggplotly(lossdata_lollipop_plot,tooltip = "text")


```
  
  
## Crosstables
Crosstables convey the same information about absolute as well as relative loss but on an aggregated portfolio level. They allow us to quickly assess whether our past and current portfolio adresses the most threatened areas. The crosstables group all PAs into different "loss groups" that are displayed in the first column (0-10 hectare, 10-100 hectare, etc.). The second column shows how many PAs fall into each group (in%) from the whole universe of PAs in partnerning countries. The third column shows how many PAs from the KfW portfolio fall into each group. Desirable would be a overrepresentation in high loss areas by KfW. Both, absolute losses (first table) and relative losses (second table) are given below. 

```{r filter_data_crosstables, message = FALSE, warning = FALSE, include=FALSE}
## Crosstables
# filter data for seperate display of jitters
wdpa_allPAs_lossdata$treated<-
  ifelse(is.na(wdpa_allPAs_lossdata$bmz_n_1)==TRUE,0,1)

wdpa_allPAs_lossdata_treated<-
  wdpa_allPAs_lossdata %>%
  filter(treated==1)
```


```{r cross_tables_loss, message = FALSE, warning = FALSE, echo=FALSE, results='asis'}
# ----- crosstables for relative and absolute loss -----
library("knitr")
# create absolute loss table
loss_data_summarycats_absolute <-
  wdpa_allPAs_lossdata %>%
  filter(treated == 0) %>%
  {
    round(table(.$absolute_loss_categorized) / nrow(.) * 100, digits = 2)
  } %>%
  as.data.frame() %>%
  rename("Absolute Forst Loss (ha)" = Var1, "% of all PAs" = Freq) 

loss_data_summarycats_absolute$`% of KfW Supported PAs` <-
  wdpa_allPAs_lossdata %>%
  filter(treated == 1) %>%
  {
    round(table(.$absolute_loss_categorized) / nrow(.) * 100, digits = 2)
  } %>%
  as.data.frame() %>%
  rename("Absolute Forst Loss (ha)" = Var1,
         "% of KfW Supported PAs" = Freq) %>%
  pull("% of KfW Supported PAs")

kable(loss_data_summarycats_absolute)

# create relative loss table
loss_data_summarycats_relative <-
  wdpa_allPAs_lossdata %>%
  filter(treated == 0) %>%
  {
    round(table(.$relative_loss_categorized) / nrow(.) * 100, digits = 2)
  } %>%
  as.data.frame() %>%
  rename("Relative Forst Loss" = Var1, "% of all PAs" = Freq) 

loss_data_summarycats_relative$`% of KfW Supported PAs` <-
  wdpa_allPAs_lossdata %>%
  filter(treated == 1) %>%
  {
    round(table(.$relative_loss_categorized) / nrow(.) * 100, digits = 2)
  } %>%
  as.data.frame() %>%
  rename("Relative Forst Loss (ha)" = Var1,
         "% of KfW Supported PAs" = Freq) %>%
  pull("% of KfW Supported PAs")

kable(loss_data_summarycats_relative)


```
  
  
  
## Jitter Plots
Jitter Plots are a different way of displaying absolut and relative loss data and the current portfolio. Data is displayed in the plot on the country level. Different colors correspond to different countries. Country names are given with their abbreviations. 
The advantage of Jitter Plots  is that individual data-points (PAs) can be identified by clicking on them in the figure. Jitter Plots allow to visualize the (highly right-skewed) data distribution in an intuitive way and compare KfWs portfolio or a suggested set of new PAs to the whole country universe.

```{r jitter_plot, message = FALSE, warning = FALSE,fig.width=10, echo=FALSE}
# ----- jitter plots absolute loss -----
# create plot
lossdata_absolute_jitter_plot <-
  wdpa_allPAs_lossdata %>%
  filter(treated == 0) %>%
  ggplot() +
  geom_jitter(
    aes(
    x = ISO3,
    y = absolute_loss,
    color = factor(ISO3),
    text = paste(
      "Area Name: ", NAME,
      "\nAbsolute forest loss: ", round(absolute_loss,digits = 0), " ha",
      "\nRelative forest loss: ", round(relative_loss,digits = 2)*100, " %",
      "\nBMZ-Number: ", bmz_n_1)
    ),
    alpha = .8) +
  geom_jitter(
    data = wdpa_allPAs_lossdata_treated,
    aes(x = ISO3, 
        y = absolute_loss,
        text = paste(
          "Area Name: ", NAME,
          "\nAbsolute forest loss: ", round(absolute_loss,digits = 0), " ha",
          "\nRelative forest loss: ", round(relative_loss,digits = 2)*100, " %",
          "\nBMZ-Number: ", bmz_n_1)),
    color = "black",
    alpha = .8
  )+
  labs(color="Country",x="")+
  scale_y_continuous(labels=function(x) format(x, big.mark = ",", scientific = FALSE),name="Forest Loss in ha")+
  theme_classic()

ggplotly(lossdata_absolute_jitter_plot,tooltip = "text")

# ----- jitter plots relative loss -----
# create plot
lossdata_relative_jitter_plot <-
  wdpa_allPAs_lossdata %>%
  filter(treated == 0) %>%
  ggplot() +
  geom_jitter(
    aes(
      x = ISO3,
      y = relative_loss,
      color = factor(ISO3),
      text = paste(
        "Area Name: ", NAME,
        "\nAbsolute forest loss: ", round(absolute_loss,digits = 0), " ha",
        "\nRelative forest loss: ", round(relative_loss,digits = 2)*100, " %",
        "\nBMZ-Number: ", bmz_n_1)
    ),
    alpha = .8) +
  geom_jitter(
    data = wdpa_allPAs_lossdata_treated,
    aes(x = ISO3, 
        y = relative_loss,
        text = paste(
          "Area Name: ", NAME,
          "\nAbsolute forest loss: ", round(absolute_loss,digits = 0), " ha",
          "\nRelative forest loss: ", round(relative_loss,digits = 2)*100, " %",
          "\nBMZ-Number: ", bmz_n_1)),
    color = "black",
    alpha = .8
  )+
  labs(color="Country",x="")+
  scale_y_continuous(labels=function(x) format(x, big.mark = ",", scientific = FALSE),name="Forest Loss in ha")+
  theme_classic()

ggplotly(lossdata_relative_jitter_plot,tooltip = "text")

```
  
  
## Lineplots
Lineplots can be used to assess the time-trend of one or several PAs. Their advantage is to display the trend data more detailed which allows users to judge whether a trend is increasing or decreasing and whether it is more likely due to natural phenomena (huge spikes) or anthropogenic causes (steaty trends). The downside of lineplots is that they are not able to compare a large set of PAs because the plots will get fairly messy if too many individual lines are displayed. This is especially troublesome if the trends of different displayed PAs differs considerably (e.g. high loss PAs vs. low loss PAs.). The following graph therefore only displays the data for a selected country. You can isolate single trends by double-clicking on a PA name. Single clicks on additional areas will allow to do a one by one comparison. 

```{r line_plots, message = FALSE, warning = FALSE,fig.width=10, echo=FALSE}
# ----- create database -----
## loss statistics
gfw_lossstats2<-
  read_csv("../../datalake/mapme.protectedareas/output/polygon/global_forest_watch/zonal_statistics_allPAs_long_temporal.csv",
           col_types = c("cfid"))

## filter for area only
gfw_lossstats2 <-
  gfw_lossstats2 %>%
  filter(name == "loss")

gfw_lossstats_lineplot <-
  merge(gfw_lossstats2,
        st_drop_geometry(wdpa_kfw),
        by.x = "WDPA_PID",
        by.y = "WDPA_PID")

gfw_lossstats_lineplot_all <-
  merge(gfw_lossstats2,
        st_drop_geometry(wdpa_allPAs),
        by.x = "WDPA_PID",
        by.y = "WDPA_PID")


# create data
gfw_lossstats_lineplot<-
  gfw_lossstats_lineplot %>% 
  select(WDPA_PID,name,year,value,bmz_n_1,ISO3,NAME)


gfw_lossstats_lineplot_all<-
  gfw_lossstats_lineplot_all %>% 
  select(WDPA_PID,name,year,value,bmz_n_1,ISO3,NAME)

# ----- lineplot with colors -----
# create labels for plot
gfw_lossstats_lineplot_all$bmz_n_1_labeled <-
  ifelse(
    is.na(gfw_lossstats_lineplot_all$bmz_n_1),
    "Not Supported",
    gfw_lossstats_lineplot_all$bmz_n_1
  )
# create color pallete for fill
palette_length <-
  gfw_lossstats_lineplot_all %>%
  filter(ISO3 == "GTM") %>%
  {
    unique(.$bmz_n_1_labeled)
  } %>%
  length()

palette_lineplot<-
  c(brewer.pal(palette_length-1,"Set1"),"grey80")

# create lineplot
lossdata_absolute_lineplot <-
  gfw_lossstats_lineplot_all %>%
  filter(ISO3 == "GTM") %>%
  ggplot() +
  geom_line(aes(x=year,
                y=value,
                group=factor(WDPA_PID),color=factor(bmz_n_1_labeled)) ,alpha=0.4)+
  labs(x="",title="Forest Loss in PAs in Guatemala")+
  scale_y_continuous(labels=function(x) format(x, big.mark = ",", scientific = FALSE),name="Forest Loss in ha")+
  scale_color_manual(name = "Project Number",values=palette_lineplot)+
  theme_classic()

ggplotly(lossdata_absolute_lineplot)

```

## Interpretation & Discussion

As outline above the GFW methodology defines forest cover loss as a _"as a stand-replacement disturbance, or a change from a forest to non-forest state."_. Loss can either be the result of human activities or natural factors such as droughts, forest fires or hurricanes, amongst others. Thus, the data currently does not allow the differentiate permanent loss and conversion from temporary loss due to natural factors. **A more in-depth analysis of pre-selected areas is therefore recommended when using the current data.**  An illustrative example for this (in the data) is Reserva Biológica Indio Maíz in Nicaragua which was hardly hit by a hurricane in 2016 and experienced subsequent forest fires in 2017. GFW displays this loss of nearly 117,000 hectares (see also interactive map) although natural regrowth of the forests is most likely happening at this point in time. 

GFW data can be used to get an assessment of old-growth and primary forests as well as associated carbon emissions. It can indicate which areas had been highly threatened either due to natural causes (hurricanes/fires) or human causes (deforestation/logging/degradation and conversion to agriculture/silviculture). Especially useful in this context is to have a look at the whole trend from 2001 to 2020. Huge spikes in the individual trend data might indicate natural causes such as fires and hurricanes. A quick search with google always helps to confirm this hypothesis or you might look at Hurricane data e.g. from NOAA (there is a link in the interactive report). More continuous growth in forest loss is probably due to the conversion of natural forests for agricultural purposes. Again you might look at our interactive map in the link above and activate the satellite layer to see what might be happening below the detected loss. GFW does not (currently) allow to detect regrowth and regeneration but it is planned to provide that feature soon.  

## Suggested Improvements
* Analyze subsequent land-cover in loss-areas. This would be possible e.g. by using Landcover data from the ESA Copernicus mission. It would allow to seperate permanent conversion of forest area e.g. for agricultural purposes from natural losses e.g. due to hurricanes. 
* Aggregate existing loss data on a finer spatial grid or create heatmaps. This would allow to better visualize highly threatened  spots on the map. This could be especially relevant to quickly identify high pressure zones in large protected areas. 
* Analyse bufferzones: This would allow to detect areas with high forest loss dynamics in its sourroundings. 

[comment]: <> (Literature)

[^1]: UNEP-WCMC and IUCN (2022), Protected Planet: The World Database on Protected Areas (WDPA) and World Database on Other Effective Area-based Conservation Measures (WD-OECM) [Online], February 2022, Cambridge, UK: UNEP-WCMC and IUCN. Available at: www.protectedplanet.net.

[^2]: "Hansen, M. C., P. V. Potapov, R. Moore, M. Hancher, S. A. Turubanova, A. Tyukavina, D. Thau, S. V. Stehman, S. J. Goetz, T. R. Loveland, A. Kommareddy, A. Egorov, L. Chini, C. O. Justice, and J. R. G. Townshend. 2013. “High-Resolution Global Maps of 21st-Century Forest Cover Change.” Science 342 (15 November): 850–53. Data available on-line from: http://earthenginepartners.appspot.com/science-2013-global-forest."
