---
title: "PAs and Forest Cover Loss - Central America"
author: "Om Prakash Bhandari (Author), Johannes Schielein (Review)"
date: "1/23/2022"
output: workflowr::wflow_html
---

```{r setup, message = FALSE,warning = FALSE,include = FALSE}
# load relevant libraries
library("tidyverse")
library("ggplot2")
library("sf")
library("leaflet")
library("leaflet.extras")
library("leaflet.extras2")
library("ggsci")
library("scales")
library("htmltools")
library("RColorBrewer")
library("plotly")
library("rmarkdown")
```

```{r database_creation, message = FALSE, warning = FALSE, include = FALSE}

##  Protected areas
uicn_kfw<-
  read_sf("~/shared/Om/test/uicn/uicn_areas.gpkg")

## create column for area coloring
uicn_kfw$REP_AREA_cat<-
  cut(uicn_kfw$REP_AREA,
      c(0,1000,5000,10000,20000,max(uicn_kfw$REP_AREA)),
      c("< 1,000 sqkm","1,001-5,000 sqkm","5,001-10,000 sqkm","10,001-20,000 sqkm",paste("20,001-",max(uicn_kfw$REP_AREA)," sqkm",sep="")))

```


```{r color_pallettes-creation, message = FALSE, warning = FALSE, include = FALSE}

## Create Color Pals for the plot data
# create colorramp function for area
pal_area <- colorFactor(
  palette = pal_npg("nrc", alpha = 0.7)(length(unique(uicn_kfw$REP_AREA_cat))),
  domain = uicn_kfw$REP_AREA_cat
)

# create colorramp function for country
colourCount = length(unique(uicn_kfw$ISO3))
getPalette = colorRampPalette(brewer.pal(9, "Set1"))

pal_country <- colorFactor(
  palette = getPalette(colourCount),
  domain = uicn_kfw$ISO3
)

```

## Interactive Map

We checked the [Protected Planet](https://www.protectedplanet.net/en/search-areas?geo_type=site) website using the name of the protected areas provided in the excel sheet and found corresponding WDPAIDs for the 19 of them whereas remaining two were either not accurate in the website or the naming convention was not clear. These two includes `EL Imposible-San Benito` and `El Imposible-El Balsamo`, however there is a protected area named `El Imposible` with WDPAID `12494` and reported area `17.65 sqkm`. But we didn't consider this polygon in our result here.

We processed the available 19 polygons using package `mapme.forest` which has access to the raster datasets from Global Forest Watch - which provides high quality datasets to compute the forest cover area and forest cover loss area for the year 2000 to 2020. The map below is an interactive leaflet map where you can toggle on and off the provided layers. 

```{r projectmap, echo=FALSE, warning=FALSE,  fig.width=8}

## Create map
# uicn_kfw_treatment_centroid<-st_transform(st_centroid(uicn_kfw_treatment_centroid),crs = 4326)
my_map <-
  leaflet() %>%
  # add external map providers
  addTiles(group = "OpenStreetMap") %>%
  addProviderTiles(providers$CartoDB.Positron, group="CartoDB") %>%
  addProviderTiles(providers$Esri.WorldImagery, group="Satellite") %>%
  addProviderTiles(providers$Esri.WorldShadedRelief, group="Topography") %>%
  addProviderTiles(providers$NASAGIBS.ViirsEarthAtNight2012, group="Nightlights") %>%
  addTiles("https://tiles.globalforestwatch.org/umd_tree_cover_loss/latest/dynamic/{z}/{x}/{y}.png",
           group="Forest Cover Loss (2001-2020)",
           #options=tileOptions(opacity = 0.7),
           attribution = "Hansen, M. C., P. V. Potapov, R. Moore, M. Hancher, S. A. Turubanova, A. Tyukavina, D. Thau, S. V. Stehman, S. J. Goetz, T. R. Loveland, A. Kommareddy, A. Egorov, L. Chini, C. O. Justice, and J. R. G. Townshend. 2013. “High-Resolution Global Maps of 21st-Century Forest Cover Change.” Science 342 (15 November): 850–53. Data available on-line from: http://earthenginepartners.appspot.com/science-2013-global-forest.")%>%
  # add own data
  addPolygons(data = uicn_kfw,opacity = 0.9,color = "orange", group = "PA Boundaries (all years)",label = ~htmlEscape(NAME),weight = 1)%>%
  addPolygons(data = uicn_kfw,opacity = 1,color = ~pal_area(REP_AREA_cat), group = "PA Area Size",label = ~htmlEscape(REP_AREA),weight = 1)%>%
  addPolygons(data = uicn_kfw,opacity = 1,color = ~pal_country(ISO3), group = "Country",label = ~htmlEscape(ISO3),weight = 1)%>%

  # fullscreen control
  addFullscreenControl() %>%
  # add legent for area
  addLegend("bottomright",
            data = uicn_kfw,
            pal = pal_area,
            values = ~REP_AREA_cat,
            title = "Total Reported Area",
            opacity = 1,
            group = "PA Area Size") %>% 
 
  addLegend("bottomright",
            data = uicn_kfw,
            pal = pal_country,
            values = ~ISO3,
            title = "Country",
            opacity = 1,
            group = "Country") %>% 
  # add layers control to define which data is shown or ommited in default view
  addLayersControl(
    baseGroups = c("CartoDB","OpenStreetMap","Satellite","Topography","Nightlights"), #"Toner",,"Regional Primary Forests (2001)"
    overlayGroups = c("PA Boundaries (all years)","Country","PA Area Size","Forest Cover Loss (2001-2020)"),
    options = layersControlOptions(collapsed = FALSE)) %>%
  # ommit certain layers
  hideGroup(group = c("Country","PA Area Size","Forest Cover Loss (2001-2020)"))

my_map
```

```{r prepData, echo=FALSE, warning=FALSE, include = FALSE}

# load uicn
uicn_wide <- read.csv("~/shared/Om/test/uicn/gfw_area_loss-wide.csv")
# subset only loss columns
uicn_wide_subset <- uicn_wide %>%
  select(c(1, 24:43))
# add column called ISO3
uicn_wide_subset$ISO3 <- c("GTM", "GTM", "GTM", "HND", "HND", "BLZ", "BLZ", "SLV", "SLV", "HND",
                           "HND", "NIC", "NIC", "CRI", "CRI", "CRI", "CRI", "PAN", "PAN")
# add column called ID as character
uicn_wide_subset$WDPA_ID <- c("12555", "12564", "67744", "23306", "555697537", "220100", "902744", "555592565",
                         "107428", "30625", "36053", "900699", "30628", "12493", "30599", "167",
                         "2553", "198343", "68135")
# remove wdpaid
uicn_wide_subset$WDPAID <- NULL
# add name
uicn_wide_subset$Name <- 
  c("Reserva Protectora de Manantiales Cerro San Gil", "Refugio de Vida Silvestre Punta de Manabique", "Reserva de la Biósfera Sierra de las Minas", "Parque Nacional Cusuco", "Parque Nacional Cuyamel-Omoa", "Reserva Marina Port Honduras",  "Parque Nacional Sarstún Temash", "Complejo Barra de Santiago", "Manglar Bola de Monte", "Reserva Biológica Laguna Karataska", "Parque Nacional Río Kruta", "Reserva Biológica Marina Cayos Miskitos y Franja Costera", "Reserva Biológica Indio Maíz", "Refugio de Vida Silvestre Barra del Colorado", "Parque Nacional Tortuguero (terrestres)",  "Parque Nacional Tortuguero (Marino)", "Parque Internacional La Amistad Costa Rica",  "Parque Internacional La Amistad Panamá", "Humedal de importancia internacional San San Pond Sak")

### All ISO3s
# pivot to long
uicn_long <- pivot_longer(uicn_wide_subset,
                          cols = c("loss_2001", "loss_2002", "loss_2003", "loss_2004", "loss_2005", "loss_2006", "loss_2007",
                                   "loss_2008", "loss_2009", "loss_2010", "loss_2011", "loss_2012", "loss_2013", "loss_2014",
                                   "loss_2015", "loss_2016", "loss_2017", "loss_2018", "loss_2019", "loss_2020"))
# separate into name and year
uicn_long.final <- uicn_long %>%
  separate(name, c("name", "year"))
```

## Interactive Lineplots

Now, we are going to visualize the results in the form of interactive lineplots.

#### Name of the Protected Areas

Though it doesn't look elegant to visualize result of all the 19 polygons in a single plot as it might be confusing for the readers, but the level of interaction provided by the `plotly` makes it easier to visualize the changes throughout the years. We can hover around the line of interest to see the actual value over the years. Also, we can zoom and pan around the graph area. However, the easiest way is to 
double-click on any one of the Protected Area's name - which in return gives us only that lineplot which we chose. We can also select multiple lines as we want and the graph will show us the plot of only the lines we selected.

Note: the area is in `hectare`.

```{r vis1, echo=FALSE, warning=FALSE, fig.width=12}

fig1 <- plot_ly(uicn_long.final, x = ~year, y = ~value,
                name = ~Name, color = ~Name, type = 'scatter', mode = 'lines+markers') %>%
  layout(title="Forest Loss Area (2001-2020) - Protected Areas")
fig1
```

#### Country ISO code

Here is the line plot of the absolute forest cover loss area (countrywise) as shown in the legend:

Note: the area is in `hectare`.

```{r vis2, echo=FALSE, warning=FALSE, fig.width=12}

fig2 <- plot_ly(uicn_long.final, x = ~year, y = ~value,
                name = ~ISO3, color = ~ISO3, type = 'scatter', mode = 'lines+markers') %>%
  layout(title="Forest Loss Area (2001-2020) - Countries")
fig2
```

## Results

Finally, here is the table containing results for the forest cover area and loss area for the year 2000 - 2020. 

`Note: area in hectare`

```{r result, echo=FALSE, warning=FALSE}

uicn_wide$Name <- 
  c("Reserva Protectora de Manantiales Cerro San Gil", "Refugio de Vida Silvestre Punta de Manabique", "Reserva de la Biósfera Sierra de las Minas", "Parque Nacional Cusuco", "Parque Nacional Cuyamel-Omoa", "Reserva Marina Port Honduras",  "Parque Nacional Sarstún Temash", "Complejo Barra de Santiago", "Manglar Bola de Monte", "Reserva Biológica Laguna Karataska", "Parque Nacional Río Kruta", "Reserva Biológica Marina Cayos Miskitos y Franja Costera", "Reserva Biológica Indio Maíz", "Refugio de Vida Silvestre Barra del Colorado", "Parque Nacional Tortuguero (terrestres)",  "Parque Nacional Tortuguero (Marino)", "Parque Internacional La Amistad Costa Rica",  "Parque Internacional La Amistad Panamá", "Humedal de importancia internacional San San Pond Sak")

uicn <- uicn_wide[,c(1,44,2:43)]

paged_table(uicn)
```

### Appendix

Lineplot with respect to WDPA IDs.

```{r allPlots, echo=FALSE, warning=FALSE, fig.width=12}

figx <- plot_ly(uicn_long.final, x = ~year, y = ~value,
                name = ~WDPA_ID, color = ~WDPA_ID, type = 'scatter', mode = 'lines+markers') %>%
  layout(title="Forest Loss Area (2001-2020) - WDPA IDs")
figx
```
